---
title: "Large language models and human crowds make the same prediction errors — analysis and verification report"
author: "Live Leonhardsen Wilhelmsen, Roya Esfandiari & Anton Gollwitzer"
date: "`r format(Sys.Date(), '%e %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    code_folding: hide
    theme: default
params:
  render_figures: true
  data_dir: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE)
library(dplyr); library(tidyr); library(purrr); library(stringr)

## Data location. All paths are resolved relative to this document's own
## folder (knitr's working directory), so the document runs from the
## repository or as a standalone copy. The deposited comparison-level tables
## (comparisons_primary.csv, comparisons_all_models.csv,
## comparisons_archive2.csv) are searched for in: params$data_dir if set, the
## repository layout (../data/derived or data/derived), a folder named
## "Original data" next to or one level above the document, and the
## document's own folder. The raw Code Ocean capsule files (9843791) are
## searched for in the same places; they are optional — when present, the
## rebuild-and-verify chunks below also run.
find_dir <- function(marker) {
  cand <- c(params$data_dir,
            "../data/derived", "data/derived",
            "../data/raw", "data/raw",
            "Original data", "../Original data", ".")
  for (d in cand[!vapply(cand, is.null, logical(1))])
    if (file.exists(file.path(d, marker))) return(normalizePath(d))
  NULL
}
derived_dir <- find_dir("comparisons_primary.csv")
raw_dir     <- find_dir("rct_responses.RDS")
if (is.null(derived_dir)) stop(
  "The deposited comparison-level tables were not found.\n",
  "Place comparisons_primary.csv, comparisons_all_models.csv and ",
  "comparisons_archive2.csv\n",
  "either in the repository's data/derived/ folder, in a folder named ",
  "'Original data'\nnext to this document, or in the document's own folder — ",
  "or set params$data_dir\nto the folder that contains them, e.g.\n",
  '  rmarkdown::render("analysis_report.Rmd", ',
  'params = list(data_dir = "C:/path/to/Original data"))')
capsule_present <- !is.null(raw_dir)

## Reading the deposited tables. A few condition names contain typographic
## quotes. When a comparison table is written out by base `write.csv()` on a
## machine whose native encoding cannot represent them (a non-UTF-8 Windows or
## C locale), R replaces each such character by its escape, e.g. "<U+201C>",
## and the escape is what ends up in the file. Reading such a copy back would
## give condition names that no longer match the ones rebuilt from the raw
## data, so the keys are normalised on read: any "<U+XXXX>" escape is turned
## back into the character it stands for, and the result is marked UTF-8. On a
## correctly written copy this is a no-op.
unescape_ucode <- function(x) {
  if (!is.character(x)) return(x)
  hits <- gregexpr("<U\\+[0-9A-Fa-f]{4,6}>", x)
  regmatches(x, hits) <- lapply(regmatches(x, hits), function(esc)
    if (!length(esc)) esc else
      intToUtf8(strtoi(substr(esc, 4L, nchar(esc) - 1L), 16L), multiple = TRUE))
  enc2utf8(x)
}
read_deposited <- function(path) {
  read.csv(path, check.names = FALSE) %>%
    mutate(across(where(is.character), unescape_ucode))
}
```

# About this document

This report accompanies our Matters Arising on Ashokkumar, Hewitt, Ghezae &
Willer, *Large language models can predict the results of social science
experiments* (Nature, 2026), https://doi.org/10.1038/s41586-026-10742-x. It
computes every quantity reported in the commentary and its Supplementary
Information, in the order of the Supplementary Information, and ends with a
**verification ledger**: a table listing each number quoted in the manuscript
next to the value computed in this render. The document refuses to knit if any
quoted value fails to reproduce.

Code is folded by default; the *Code* buttons unfold any chunk. All random
components use fixed seeds, set at the top of the chunk that uses them.

**Three ways to use this document.**

1. *Read it.* The rendered HTML contains every number and figure; no software
   is required.
2. *Re-render it from the deposited data.* The comparison-level tables in
   `data/derived/` are included in this repository (CC BY 4.0). Knitting the
   document from them reproduces every reported value in roughly ten minutes:
   `rmarkdown::render("docs/analysis_report.Rmd")`.
3. *Re-render it from the raw data.* The raw files are **not redistributed
   here**; they are available in Code Ocean capsule 9843791, released with the
   original Article. With those files placed in `data/raw/` (see
   `data/raw/README.md`), the additional chunks below rebuild the
   comparison-level tables from scratch and verify that they are identical to
   the deposit.

The document locates its data automatically: the repository layout
(`data/derived/`, `data/raw/`), a folder named `Original data` next to the
document, or the document's own folder — and `params$data_dir` overrides the
search, so a standalone copy of this file knits anywhere as long as the three
comparison-level tables are nearby.

The numbered scripts in `R/` run the same analyses as a pipeline
(`Rscript run_all.R`); this document is the self-contained, readable version.
In this render, the capsule files are
`r if (capsule_present) "present, so the rebuild-and-verify chunks run" else "not present, so the rebuild-and-verify chunks are skipped and the deposited tables are used directly"`.

# Data (Supplementary Methods 1.1)

The deposited comparison-level tables are the working data for everything
below: one row per condition comparison, with the observed effect, its
standard error, and each predictor's raw and recalibrated standardized errors.

```{r derived-data}
cat("Comparison-level tables found in:", derived_dir, "\n")
wide <- read_deposited(file.path(derived_dir, "comparisons_primary.csv"))
all_models <- read_deposited(file.path(derived_dir, "comparisons_all_models.csv"))

c(comparisons         = nrow(wide),
  studies             = n_distinct(wide$study),
  study_outcome_cells = wide %>% distinct(study, outcome.name) %>% nrow())
```

The raw-data counts reported in Supplementary Methods 1.1 require the capsule
files and are computed here when those are present.

```{r capsule-counts, eval=capsule_present}
rct <- readRDS(file.path(raw_dir, "rct_responses.RDS"))
llm <- readRDS(file.path(raw_dir, "llm_responses.RDS"))
fc  <- readRDS(file.path(raw_dir, "forecasting_responses.RDS"))

c(rct_responses      = sum(map_int(rct$data, nrow)),
  answers_nonmissing = sum(map_int(rct$data, ~sum(!is.na(.x$y)))),
  studies_raw        = n_distinct(rct$study),
  study_outcome_cells_raw = nrow(rct),
  llm_rows           = nrow(llm),
  llm_rows_per_model = nrow(llm) / n_distinct(llm$model),
  forecast_rows      = nrow(fc),
  forecasters        = n_distinct(fc$PROLIFIC_PID))

## the remaining descriptive counts quoted in Supplementary Methods 1.1
per_cell       <- fc %>% count(study, outcome.name, condition.name)
per_forecaster <- fc %>% group_by(PROLIFIC_PID) %>%
  summarise(conditions = n_distinct(study, outcome.name, condition.name),
            studies    = n_distinct(study), .groups = "drop")
c(median_forecasts_per_cell    = median(per_cell$n),
  median_conditions_per_person = median(per_forecaster$conditions),
  forecasters_in_multiple_experiments = sum(per_forecaster$studies > 1),
  prompt_cells_240   = llm %>% count(model, study, outcome.name, condition.name) %>%
                         filter(model == "gpt-4", n == 240) %>% nrow(),
  possible_pairs     = rct %>% pmap_int(function(data, ...) {
                         k <- n_distinct(data$condition.name)
                         as.integer(k * (k - 1) / 2) }) %>% sum())
```

# From raw data to the comparison table (Supplementary Methods 1.2–1.3)

The chunks in this section document the full construction: observed effects
from the participant-level responses, predicted effects from the forecast
condition means, one row per unordered comparison, and the authors'
leave-one-study-out linear recalibration. They execute when the capsule files
are present and finish by verifying that the rebuilt table matches the
deposited `comparisons_primary.csv` exactly. The analyses in the rest of the
document run from the deposited table either way.

**Observed effects.** Within each study and outcome, responses are rescaled to
0–1 using the bounds of the outcome scale, and every pairwise difference
between conditions is estimated by ordinary least squares, retaining the
estimate and its standard error. Condition names are recoded to plain ASCII
before the regression and mapped back afterwards: some names contain
typographic quotes, which R on a non-UTF-8 locale silently rewrites inside
coefficient names, and the recode makes the construction identical on any
machine.

```{r observed-effects, eval=capsule_present, cache=TRUE}
rct_effects <- rct %>%
  pmap_dfr(function(study, outcome.name, outcome.min, outcome.max, data, ...) {
    d <- data %>% ungroup() %>%
      mutate(y = (as.numeric(y) - outcome.min) / (outcome.max - outcome.min))
    conds <- unique(d$condition.name)
    code  <- setNames(paste0("k", seq_along(conds)), conds)
    d$cc  <- code[d$condition.name]
    map_dfr(conds, function(ref) {
      fit <- lm(y ~ cc, data = d %>% mutate(cc = forcats::fct_relevel(cc, code[[ref]])))
      broom::tidy(fit) %>%
        transmute(cc = str_match(term, "^cc(k\\d+)$")[, 2], estimate, std.error) %>%
        filter(!is.na(cc)) %>%
        mutate(condition.name = names(code)[match(cc, code)],
               study = study, outcome.name = outcome.name,
               reference_condition = ref, .before = 1) %>%
        select(-cc)
    })
  })
```

**Predicted effects.** For each language model, the predicted mean of a
condition is the weighted average of its prompt-level responses, using the
authors' demographic weights; reversed answer scales are un-flipped first,
following their code. For the human crowd, the predicted mean is the
unweighted average of the individual forecasts. A predicted effect is the
difference between the predicted means of the two conditions.

```{r predictions, eval=capsule_present, cache=TRUE}
llm_means <- llm %>%
  mutate(expectation = ifelse(scale_flip, outcome_scale_max - (expectation - outcome_scale_min),
                              expectation),
         y = (expectation - outcome_scale_min) / (outcome_scale_max - outcome_scale_min)) %>%
  filter(!is.na(y)) %>%
  group_by(model, study, outcome.name, condition.name) %>%
  summarise(pred_mean = weighted.mean(y, w = weight), .groups = "drop")

fc_means <- fc %>%
  mutate(y = value / 100) %>%
  semi_join(distinct(llm, study, outcome.name), by = c("study", "outcome.name")) %>%
  group_by(study, outcome.name, condition.name) %>%
  summarise(pred_mean = mean(y), .groups = "drop") %>%
  mutate(model = "human")

pred_contrasts <- bind_rows(llm_means, fc_means) %>%
  group_by(model, study, outcome.name) %>%
  group_modify(function(d, k) {
    crossing(d %>% select(condition.name, pred_mean),
             d %>% select(reference_condition = condition.name, ref_mean = pred_mean)) %>%
      filter(condition.name != reference_condition) %>%
      transmute(condition.name, reference_condition, prediction = pred_mean - ref_mean)
  }) %>%
  ungroup()
```

**One row per comparison, and the combined predictor.** Comparing A with B and
B with A is the same comparison with the sign flipped, so each unordered pair
is kept once. The direction of each pair is fixed by ordering condition names
with a fixed ICU collator (English): ordering with `<` depends on the
computer's locale — in a plain C locale, R returns `NA` for names with
typographic quotes and silently drops those comparisons — whereas the ICU
ordering is identical on every machine and matches the orientation of the
deposited dataset. The authors' proposed combination of model and crowd is the
arithmetic mean of the two predictions, built the same way so it can be tested
below.

```{r comparisons, eval=capsule_present, cache=TRUE}
en <- stringi::stri_opts_collator(locale = "en_US")
comparisons_built <- rct_effects %>%
  inner_join(pred_contrasts,
             by = c("study", "outcome.name", "condition.name", "reference_condition")) %>%
  filter(stringi::stri_cmp_lt(reference_condition, condition.name, opts_collator = en))

combined_built <- comparisons_built %>%
  filter(model %in% c("gpt-4", "human")) %>%
  select(study, outcome.name, condition.name, reference_condition,
         estimate, std.error, model, prediction) %>%
  pivot_wider(names_from = model, values_from = prediction) %>%
  filter(!is.na(`gpt-4`), !is.na(human)) %>%
  mutate(model = "combined", prediction = (`gpt-4` + human) / 2) %>%
  select(-`gpt-4`, -human)

comparisons_built <- bind_rows(comparisons_built, combined_built)
```

**Recalibration.** Because the raw predictions systematically overstate effect
magnitudes, each predictor is recalibrated with the authors' leave-one-study-out
linear procedure, which they used for their pilot-testing analysis: for each
study, `estimate ~ prediction` is fitted on the other 69 studies and applied to
the held-out study, so no study contributes to its own recalibration. Each
prediction error is then expressed in units of the observed effect's standard
error, and a prediction is classified as a miss when it differs from the
observed effect by more than 1.96 standard errors.

```{r recalibration, eval=capsule_present, cache=TRUE}
comparisons_built <- comparisons_built %>%
  mutate(z_raw = (prediction - estimate) / std.error) %>%
  group_by(model) %>%
  group_modify(function(d, k) {
    d$pred_recal <- NA_real_
    for (s in unique(d$study)) {
      fit <- lm(estimate ~ prediction, data = d %>% filter(study != s))
      d$pred_recal[d$study == s] <- predict(fit, newdata = d %>% filter(study == s))
    }
    d
  }) %>%
  ungroup() %>%
  mutate(z_recal = (pred_recal - estimate) / std.error)

wide_built <- comparisons_built %>%
  filter(model %in% c("gpt-4", "human", "combined")) %>%
  select(study, outcome.name, reference_condition, condition.name,
         estimate, std.error, model, z_recal) %>%
  pivot_wider(names_from = model, values_from = z_recal) %>%
  filter(!is.na(`gpt-4`), !is.na(human))
```

```{r verify-deposit, eval=capsule_present}
## the deposited table is identical to the rebuild
m <- inner_join(
  wide_built %>% transmute(study, outcome.name, reference_condition, condition.name,
                           zg_built = `gpt-4`, zh_built = human, zc_built = combined),
  wide %>% select(study, outcome.name, reference_condition, condition.name, zg, zh, zc),
  by = c("study", "outcome.name", "reference_condition", "condition.name"))
stopifnot(nrow(m) == nrow(wide), nrow(m) == nrow(wide_built),
          max(abs(m$zg_built - m$zg)) < 1e-8,
          max(abs(m$zh_built - m$zh)) < 1e-8,
          max(abs(m$zc_built - m$zc)) < 1e-8)
cat("Rebuilt table matches the deposited comparisons_primary.csv:",
    nrow(m), "comparisons, maximum |difference| <", 1e-8, "\n")
```

# Miss rates and recalibration summary (Supplementary Methods 1.3)

```{r miss-rates}
miss_g <- abs(wide$zg) > 1.96
miss_h <- abs(wide$zh) > 1.96

## pooled recalibration lines (the leave-one-study-out fits vary only slightly
## around these); pg and ph are the unrecalibrated predicted effects
recal_lines <- bind_rows(
  broom::tidy(lm(estimate ~ pg, data = wide)) %>% mutate(model = "gpt-4"),
  broom::tidy(lm(estimate ~ ph, data = wide)) %>% mutate(model = "human")) %>%
  transmute(model, term = ifelse(grepl("Intercept", term), "intercept", "slope"),
            estimate = round(estimate, 3))
recal_lines

c(miss_gpt4_raw     = round(mean(abs(wide$zg_raw) > 1.96), 3),
  miss_crowd_raw    = round(mean(abs(wide$zh_raw) > 1.96), 3),
  miss_gpt4_recal   = round(mean(miss_g), 3),
  miss_crowd_recal  = round(mean(miss_h), 3))
```

# The headline result (main text)

Across the 1,678 comparisons, the two predictors' misses coincide — and point
the same way — far more often than unrelated predictors would.

```{r core}
joint <- miss_g & miss_h
same  <- joint & (sign(wide$zg) == sign(wide$zh))

tibble(
  `Joint misses`      = sum(joint),
  `Joint rate`        = mean(joint),
  `Same-side misses`  = sum(same),
  `Same-side rate`    = mean(same),
  `Direction | joint` = sum(same) / sum(joint)
) %>% mutate(across(where(is.numeric), ~round(., 3))) %>% as.data.frame()
```

**The combined predictor.** The authors propose averaging model and crowd. The
average was built and recalibrated identically above, so its behaviour on the
joint misses can be read off directly: averaging repairs opposite-side misses
and leaves same-side misses in place.

```{r combined}
comb <- c(
  recovers_opposite_side = mean(abs(wide$zc[joint & !same]) <= 1.96),
  still_outside_same_side = mean(abs(wide$zc[same]) > 1.96))
round(comb, 3)
```

# The corrected null (Supplementary Methods 1.4)

Some same-direction overlap is expected even between independent predictors,
because both are scored against the same estimated effect, which carries
sampling noise of its own. Writing each standardized error as $z_i = a_i - u$,
where $a_i$ is the predictor's own deviation from the true effect and $u$ is
the estimate's standardized sampling noise — shared by both predictors, with
variance 1 by construction — the corrected null treats the two predictors as
independent while retaining the shared $u$, with each predictor's spread
calibrated so that its overall miss rate matches the observed one. Under this
null the pair of errors is bivariate normal, so every benchmark is a
one-dimensional integral and carries no simulation noise.

```{r exact-null}
null_exact <- function(miss_g, miss_h, t = 1.96) {
  sig <- function(p) t / qnorm(1 - p / 2)            # total sd yielding this miss rate
  s2g <- max(sig(miss_g)^2 - 1, 0); s2h <- max(sig(miss_h)^2 - 1, 0)
  tail_hi <- function(s2, u) if (s2 > 0) pnorm((t + u) / sqrt(s2), lower.tail = FALSE) else as.numeric(-u >  t)
  tail_lo <- function(s2, u) if (s2 > 0) pnorm((-t + u) / sqrt(s2))                    else as.numeric(-u < -t)
  f <- function(u, kind) sapply(u, function(uu) {
    gh <- tail_hi(s2g, uu); gl <- tail_lo(s2g, uu)
    hh <- tail_hi(s2h, uu); hl <- tail_lo(s2h, uu)
    v <- if (kind == "joint") (gh + gl) * (hh + hl) else gh * hh + gl * hl
    v * dnorm(uu)
  })
  joint <- integrate(f, -9, 9, kind = "joint", rel.tol = 1e-9, subdivisions = 500L)$value
  same  <- integrate(f, -9, 9, kind = "same",  rel.tol = 1e-9, subdivisions = 500L)$value
  c(joint = joint, same = same, direction = same / joint)
}
ne <- null_exact(mean(miss_g), mean(miss_h))
round(100 * ne, 2)

## the naive benchmark that ignores the shared estimate
c(naive_same_side = round(mean(miss_g) * mean(miss_h) / 2, 3), naive_direction = 0.5)
```

A direct simulation of the same null verifies the integral and provides the
calibrated spreads reused by the exceedance tests below.

```{r corrected-null, cache=TRUE}
set.seed(2)
N <- 2e6
u <- rnorm(N); bg <- rnorm(N); bh <- rnorm(N)
calib <- function(b, target)
  uniroot(function(s) mean(abs(s * b - u) > 1.96) - target, c(0.05, 20))$root
sg <- calib(bg, mean(miss_g)); sh <- calib(bh, mean(miss_h))
zgs <- sg * bg - u; zhs <- sh * bh - u
joint_s <- abs(zgs) > 1.96 & abs(zhs) > 1.96
same_s  <- joint_s & sign(zgs) == sign(zhs)
c(null_joint = round(mean(joint_s), 3),
  null_same  = round(mean(same_s), 3),
  null_direction = round(mean(sign(zgs[joint_s]) == sign(zhs[joint_s])), 3),
  E_zg_zh = round(mean(zgs * zhs), 3))   # theory: exactly Var(u) = 1 under this null
```

**Sensitivity to the normality of the null.** The predictors' real errors have
heavier tails than a normal distribution, so the two decisive benchmarks are
recomputed exactly under heavier-tailed error shapes — a Student t with 3
degrees of freedom, and the empirical shape of the errors themselves. Heavier
tails lower the benchmarks slightly, so the normal specification reported in
the manuscript is the most conservative.

```{r tail-variants, cache=TRUE}
null_exact_tail <- function(miss_g, miss_h, t = 1.96, ptail) {
  miss_of <- function(s) integrate(function(u) sapply(u, function(uu)
      (ptail((t + uu) / s) + ptail((t - uu) / s)) * dnorm(uu)), -9, 9,
      subdivisions = 500L, stop.on.error = FALSE)$value
  cal <- function(p) uniroot(function(s) miss_of(s) - p, c(0.05, 30))$root
  sg <- cal(miss_g); sh <- cal(miss_h)
  f <- function(u, kind) sapply(u, function(uu) {
    gh <- ptail((t + uu) / sg); gl <- ptail((t - uu) / sg)
    hh <- ptail((t + uu) / sh); hl <- ptail((t - uu) / sh)
    v <- if (kind == "joint") (gh + gl) * (hh + hl) else gh * hh + gl * hl
    v * dnorm(uu)
  })
  joint <- integrate(f, -9, 9, kind = "joint", subdivisions = 500L, stop.on.error = FALSE)$value
  same  <- integrate(f, -9, 9, kind = "same",  subdivisions = 500L, stop.on.error = FALSE)$value
  c(same = 100 * same, direction = 100 * same / joint)
}
t3p <- function(q) pt(q * sqrt(3), df = 3, lower.tail = FALSE)   # t(3), unit variance
eg  <- (wide$zg - mean(wide$zg)) / sd(wide$zg)                   # empirical shapes
eh  <- (wide$zh - mean(wide$zh)) / sd(wide$zh)
emp <- function(z) function(q) sapply(q, function(qq) mean(z > qq))
tails <- rbind(
  `normal`       = 100 * ne[c("same", "direction")],
  `Student t(3)` = null_exact_tail(mean(miss_g), mean(miss_h), ptail = t3p),
  `empirical`    = null_exact_tail(mean(miss_g), mean(miss_h),
                                   ptail = function(q) (emp(eg)(q) + emp(eh)(q)) / 2))
round(tails, 1)
```

# The study-specific null (Supplementary Methods 1.5)

The corrected null matches each predictor's miss rate across the full archive,
assuming every study is equally hard to predict. Studies differ in difficulty,
and in a hard study both predictors miss more often even when their errors are
unrelated. The stricter benchmark below therefore calibrates the two spreads
within each study to that study's own miss rates and pools across studies,
4,000 simulated replicates per real comparison. When a predictor's miss rate
within a study is at or below 5% — the rate attributable to sampling error
alone — the spread is set to zero, which aligns the two simulated predictors
perfectly and inflates the null: the conservative direction.

```{r study-specific-null, cache=TRUE}
set.seed(21)
POOL   <- 1e6
pool_b <- rnorm(POOL); pool_u <- rnorm(POOL)
rate_for <- function(s) mean(abs(s * pool_b - pool_u) > 1.96)
calib_s <- function(target) {
  if (target <= rate_for(0))  return(0)
  if (target >= rate_for(40)) return(40)
  uniroot(function(s) rate_for(s) - target, c(0, 40), tol = 1e-4)$root
}

REP <- 4000
ss_null <- map_dfr(unique(wide$study), function(k) {
  m  <- wide$study == k
  nk <- sum(m)
  s1 <- calib_s(mean(miss_g[m]))
  s2 <- calib_s(mean(miss_h[m]))
  uu <- rnorm(nk * REP); b1 <- rnorm(nk * REP); b2 <- rnorm(nk * REP)
  z1 <- s1 * b1 - uu; z2 <- s2 * b2 - uu
  j  <- abs(z1) > 1.96 & abs(z2) > 1.96
  tibble(n = length(z1), joint = sum(j), same = sum(j & sign(z1) == sign(z2)))
})

ss_rates <- c(joint     = sum(ss_null$joint) / sum(ss_null$n),
              same_side = sum(ss_null$same)  / sum(ss_null$n),
              direction = sum(ss_null$same)  / sum(ss_null$joint))
rbind(
  observed              = c(mean(joint), mean(same), sum(same) / sum(joint)),
  `global null`         = ne[c("joint", "same", "direction")],
  `study-specific null` = ss_rates
) %>% `colnames<-`(c("joint", "same_side", "direction")) %>% round(3)
```

The study-specific benchmark takes differences in prediction difficulty as
given; because observed study-level miss rates may already contain shared
prediction error, it can absorb part of the dependence being tested and is
therefore the more conservative of the two. Study difficulty raises the
joint-miss benchmark (14.1% to 16.0%) but not the directional ones: the
same-side benchmark barely moves and the direction benchmark falls, so the
observed 90% clears it by more.

# Uncertainty intervals (Supplementary Methods 1.6)

Comparisons from the same study are not independent — they share participants
and often a reference condition — so uncertainty is estimated by resampling
studies rather than comparisons: a nonparametric cluster bootstrap that
resamples the 70 studies with replacement 4,000 times, with percentile
intervals.

```{r bootstrap, cache=TRUE}
set.seed(3)
studies <- unique(wide$study)
boot <- replicate(4000, {
  s <- sample(studies, replace = TRUE)
  d <- map_dfr(s, ~ wide[wide$study == .x, ])
  jg <- abs(d$zg) > 1.96 & abs(d$zh) > 1.96
  sm <- jg & sign(d$zg) == sign(d$zh)
  c(joint = mean(jg), same = mean(sm), dir = sum(sm) / sum(jg))
})
boot_ci <- apply(boot, 1, quantile, c(.025, .975))
round(boot_ci, 3)
```

Conditional shares — the proportion of GPT-4's misses that another predictor
repeats, and the proportions in the secondary archive, where only 15 megastudy
families are available for resampling — carry Wilson score intervals. The
error-on-error slope carries a cluster-robust (sandwich) confidence interval.
These helpers are used throughout the remaining sections.

```{r interval-helpers}
wilson_ci <- function(k, n, level = .95) {
  if (n == 0) return(c(lo = NA_real_, hi = NA_real_))
  z <- qnorm(1 - (1 - level) / 2); p <- k / n; d <- 1 + z^2 / n
  c(lo = (p + z^2/(2*n) - z * sqrt(p*(1-p)/n + z^2/(4*n^2))) / d,
    hi = (p + z^2/(2*n) + z * sqrt(p*(1-p)/n + z^2/(4*n^2))) / d)
}

cluster_vcov <- function(fit, cluster) {
  X <- model.matrix(fit); u <- residuals(fit)
  XtXi <- solve(crossprod(X))
  meat <- Reduce(`+`, lapply(split(seq_len(nrow(X)), cluster), function(i) {
    v <- crossprod(X[i, , drop = FALSE], u[i]); tcrossprod(v)
  }))
  XtXi %*% meat %*% XtXi
}

slope_band <- function(x, y, cluster) {
  fit <- lm(y ~ x)
  V   <- cluster_vcov(fit, cluster)
  list(slope = unname(coef(fit)[2]), slope_se = sqrt(V[2, 2]))
}
```

# Exceedance tests (Supplementary Methods 1.7)

The corrected null gives the rates independent predictors are expected to
produce; the tests below ask whether sampling variation under that null could
reach the observed rates. Three complementary one-sided tests, from three
angles; they agree.

**Test 1 — Monte Carlo against the global corrected null.** 10,000 archives of
1,678 comparisons are simulated from the corrected null, using the calibrated
spreads from Supplementary Methods 1.4. The P value is the share of simulated
archives whose rate reaches the observed one; the null's 2.5th–97.5th
percentile range provides the error bars drawn on the null benchmarks in the
figures.

```{r exceedance-global, cache=TRUE}
set.seed(41)
R_NULL <- 10000
obs <- c(joint = mean(joint), same_side = mean(same), direction = sum(same) / sum(joint))

null_draw <- function(n) {
  uu <- rnorm(n)
  z1 <- sg * rnorm(n) - uu
  z2 <- sh * rnorm(n) - uu
  j  <- abs(z1) > 1.96 & abs(z2) > 1.96
  s  <- j & sign(z1) == sign(z2)
  c(joint = mean(j), same_side = mean(s),
    direction = if (sum(j)) sum(s) / sum(j) else NA_real_)
}
null_dist <- t(replicate(R_NULL, null_draw(nrow(wide))))

p_mc <- sapply(colnames(null_dist), function(k)
  (1 + sum(null_dist[, k] >= obs[[k]], na.rm = TRUE)) / (1 + R_NULL))
null_ci <- apply(null_dist, 2, quantile, c(.025, .975), na.rm = TRUE)

rbind(observed = obs, null_lo = null_ci[1, ], null_hi = null_ci[2, ], P = p_mc) %>% round(4)
```

**Test 2 — Monte Carlo against the study-specific null.** The same test
against the stricter benchmark of Supplementary Methods 1.5, with 2,000
simulated archives and the per-study spreads calibrated once and reused.

```{r exceedance-ss, cache=TRUE}
set.seed(42)
R_SS <- 2000

study_tab <- wide %>%
  group_by(study) %>%
  summarise(n = n(), mg = mean(abs(zg) > 1.96), mh = mean(abs(zh) > 1.96),
            .groups = "drop") %>%
  mutate(s1 = sapply(mg, calib_s), s2 = sapply(mh, calib_s))

ss_draw <- function() {
  tot <- sum(study_tab$n); jj <- 0L; ss <- 0L
  for (i in seq_len(nrow(study_tab))) {
    nk <- study_tab$n[i]
    uu <- rnorm(nk)
    z1 <- study_tab$s1[i] * rnorm(nk) - uu
    z2 <- study_tab$s2[i] * rnorm(nk) - uu
    j  <- abs(z1) > 1.96 & abs(z2) > 1.96
    jj <- jj + sum(j); ss <- ss + sum(j & sign(z1) == sign(z2))
  }
  c(joint = jj / tot, same_side = ss / tot,
    direction = if (jj) ss / jj else NA_real_)
}
ss_dist <- t(replicate(R_SS, ss_draw()))

p_ss  <- sapply(colnames(ss_dist), function(k)
  (1 + sum(ss_dist[, k] >= obs[[k]], na.rm = TRUE)) / (1 + R_SS))
ss_ci <- apply(ss_dist, 2, quantile, c(.025, .975), na.rm = TRUE)

rbind(observed = obs, null_lo = ss_ci[1, ], null_hi = ss_ci[2, ], P = p_ss) %>% round(4)
```

**Test 3 — study-bootstrap exceedance.** The Monte Carlo tests draw
comparisons independently, whereas real comparisons cluster within studies.
Inverting the study-level bootstrap into a test respects that structure: the
studies are resampled with replacement 4,000 times, and the P value is the
share of draws whose rate falls at or below the corrected-null benchmark.
(This is an independent bootstrap instance with its own seed, as in the
pipeline; the confidence intervals of Supplementary Methods 1.6 come from the
bootstrap above. At this resolution the smallest P values are bounded by the
number of draws.)

```{r exceedance-boot, cache=TRUE}
set.seed(43)
idx_by_study_b <- split(seq_len(nrow(wide)), wide$study)
boot_e <- replicate(4000, {
  s  <- sample(studies, replace = TRUE)
  ii <- unlist(idx_by_study_b[s], use.names = FALSE)
  j  <- joint[ii]; sm <- same[ii]
  c(joint = mean(j), same_side = mean(sm),
    direction = if (sum(j)) sum(sm) / sum(j) else NA_real_)
})
p_boot <- c(joint     = mean(boot_e["joint", ]     <= ne[["joint"]]),
            same_side = mean(boot_e["same_side", ] <= ne[["same"]]),
            direction = mean(boot_e["direction", ] <= ne[["direction"]], na.rm = TRUE))
round(p_boot, 4)
```

None of the 10,000 corrected-null archives and none of the 2,000
study-specific archives reaches the observed same-direction rate or direction
share, and for the direction share no bootstrap draw falls at or below the
null. The observed overlap is not something sampling variation under
independence produces. The assumption-free permutation test in Supplementary
Results 2.4 below makes the same point without any distributional assumptions.

# Robustness to recalibration and threshold (Supplementary Results 2.1)

The shared-error pattern is present without recalibration, under linear and
isotonic recalibration — the isotonic variant fitted with the same
leave-one-study-out scheme — and at stricter miss thresholds, in each case
against its own corrected-null benchmark computed at the same threshold.

```{r robustness, cache=TRUE}
iso_loo <- function(pred, est, study) {
  out <- rep(NA_real_, length(pred))
  for (s in unique(study)) {
    tr <- study != s
    fit <- isoreg(pred[tr], est[tr])
    f <- approxfun(sort(fit$x), fit$yf, method = "linear", rule = 2)
    out[!tr] <- f(pred[!tr])
  }
  out
}
zg_iso <- (iso_loo(wide$pg, wide$estimate, wide$study) - wide$estimate) / wide$std.error
zh_iso <- (iso_loo(wide$ph, wide$estimate, wide$study) - wide$estimate) / wide$std.error

summ <- function(a, b, thr = 1.96) {
  j <- abs(a) > thr & abs(b) > thr; s <- j & sign(a) == sign(b)
  nulls <- null_exact(mean(abs(a) > thr), mean(abs(b) > thr), t = thr)
  tibble(miss_g = mean(abs(a) > thr), miss_h = mean(abs(b) > thr),
         joint = mean(j), null_joint = nulls[["joint"]],
         same_side = mean(s), null_same = nulls[["same"]],
         direction = sum(s)/sum(j), null_direction = nulls[["direction"]])
}
rob <- bind_rows(
  `no recalibration`       = summ(wide$zg_raw, wide$zh_raw),
  `linear recalibration`   = summ(wide$zg, wide$zh),
  `isotonic recalibration` = summ(zg_iso, zh_iso),
  `linear, threshold 1.64` = summ(wide$zg, wide$zh, 1.64),
  `linear, threshold 2.58` = summ(wide$zg, wide$zh, 2.58),
  `linear, threshold 3.29` = summ(wide$zg, wide$zh, 3.29),
  .id = "analysis")
rob %>% mutate(across(where(is.numeric), ~round(., 3))) %>% as.data.frame()
```

On the original percentage-point scale, the two predictors' errors correlate
with and without recalibration:

```{r pp-corr}
c(recalibrated = round(cor(wide$zg * wide$std.error, wide$zh * wide$std.error), 2),
  uncorrected  = round(cor(wide$pg - wide$estimate, wide$ph - wide$estimate), 2))
```

# Study weighting and the authors' sampling scheme (Supplementary Results 2.2)

The 70 studies contribute between 1 and 198 comparisons each, so the headline
rates weight condition-rich studies more heavily. Giving each study equal
weight raises both rates — the reported comparison-weighted figures are the
conservative choice.

```{r study-weighted}
by_study <- wide %>%
  group_by(study) %>%
  summarise(n     = n(),
            joint = mean(abs(zg) > 1.96 & abs(zh) > 1.96),
            same  = mean(abs(zg) > 1.96 & abs(zh) > 1.96 & sign(zg) == sign(zh)),
            .groups = "drop")
c(joint_comparison_weighted = mean(joint),
  joint_study_weighted      = mean(by_study$joint),
  same_comparison_weighted  = mean(same),
  same_study_weighted       = mean(by_study$same)) %>% round(3)
range(by_study$n)
```

The Article's headline analysis instead samples one outcome and one reference
condition per study and compares every other condition with that reference,
repeated over 32 draws. Applying that scheme exactly:

A single 32-draw run is itself a random quantity, and 32 draws are too few to
pin it down: which comparisons a draw selects depends on the order in which a
study's outcomes and conditions are encountered, so the figure moves both with
the seed and with the row order of the table it is drawn from. Repeating the
whole 32-draw procedure 200 times measures that movement, and the scheme is
reported here at 500 draws, where it has settled. The 500-draw figures are
quoted to two decimal places, the precision at which they are reproducible
from any ordering of the comparison table.

```{r authors-sampling, cache=TRUE}
set.seed(22)
idx_by_study <- split(seq_len(nrow(wide)), wide$study)

one_draw <- function() {
  keep <- unlist(lapply(idx_by_study, function(ii) {
    ocs <- unique(wide$outcome.name[ii])
    oc  <- ocs[sample.int(length(ocs), 1)]
    jj  <- ii[wide$outcome.name[ii] == oc]
    cds <- unique(c(wide$reference_condition[jj], wide$condition.name[jj]))
    ref <- cds[sample.int(length(cds), 1)]
    jj[wide$reference_condition[jj] == ref | wide$condition.name[jj] == ref]
  }), use.names = FALSE)
  g <- wide$zg[keep]; h <- wide$zh[keep]
  j <- abs(g) > 1.96 & abs(h) > 1.96
  c(n = length(keep), joint = mean(j), same = mean(j & sign(g) == sign(h)))
}
draws500 <- t(replicate(500, one_draw()))

## the Article reports one 32-draw run; repeating that whole run 200 times
## shows how far such a figure travels from one run to the next
rep32 <- t(replicate(200, {
  d <- t(replicate(32, one_draw()))
  c(joint = median(d[, "joint"]), same = median(d[, "same"]))
}))

rbind(`all 1,678 comparisons`   = c(NA, mean(joint), mean(same)),
      `their scheme, 500 draws` = c(median(draws500[, "n"]),
                                    median(draws500[, "joint"]), median(draws500[, "same"]))
) %>% `colnames<-`(c("comparisons_per_draw", "joint", "same_side")) %>% round(3)

## how much a single 32-draw run moves between runs
apply(rep32, 2, function(x) c(sd = sd(x), lo = quantile(x, .025),
                              hi = quantile(x, .975))) %>% round(3)

## spread across the individual draws — a sensitivity check, not a confidence
## interval; the intervals in the manuscript come from the study bootstrap above
apply(draws500[, c("joint", "same")], 2, sd) %>% round(3)
```

Each draw retains a median of 235 unique comparisons; the Article's 469
effects count the contrasts in both orientations. The rates are unaffected by
orientation, because flipping a pair flips the sign of the observed effect and
of both predictions together, leaving every |z| and every sign agreement
unchanged.

# Study-level heterogeneity (Supplementary Results 2.3)

```{r heterogeneity}
c(studies_with_joint_miss = n_distinct(wide$study[joint]),
  studies_with_same_side  = n_distinct(wide$study[same]))

## error-on-error slope with cluster-robust (by study) confidence interval
sb1 <- slope_band(wide$zg, wide$zh, wide$study)
het_slope <- c(slope = round(sb1$slope, 2),
               ci_lo = round(sb1$slope - 1.96 * sb1$slope_se, 2),
               ci_hi = round(sb1$slope + 1.96 * sb1$slope_se, 2))
het_slope

## share of the co-error's variance lying between studies (one-way ANOVA)
w  <- wide$zg * wide$zh
gm <- mean(w)
ssb <- sum(tapply(w, wide$study, function(x) length(x) * (mean(x) - gm)^2))
varshare <- round(100 * ssb / sum((w - gm)^2), 1)
varshare

## per-study slopes, among studies with at least five comparisons
slopes <- wide %>% group_by(study) %>% filter(n() >= 5, sd(zg) > 0) %>%
  summarise(slope = coef(lm(zh ~ zg))[2], .groups = "drop")
slope_sum <- c(n_studies = nrow(slopes), median = round(median(slopes$slope), 2),
               q25 = round(unname(quantile(slopes$slope, .25)), 2),
               q75 = round(unname(quantile(slopes$slope, .75)), 2))
slope_sum
```

The figures print a cluster-bootstrap interval for the same slope alongside
the fitted line; the manuscript text quotes the cluster-robust (sandwich)
interval computed here. Both cluster on the study.

# The continuous co-error (Supplementary Results 2.4)

The findings do not depend on classifying predictions as misses. The
continuous co-error multiplies the two standardized errors, comparison by
comparison; it is positive when they point the same way. Randomly re-pairing
the two error series destroys any dependence while preserving each marginal
distribution exactly.

```{r permutation}
set.seed(1)
w_obs  <- mean(wide$zg * wide$zh)
w_perm <- replicate(2000, mean(wide$zg * sample(wide$zh)))
per_study <- wide %>% group_by(study) %>% summarise(w = mean(zg * zh), .groups = "drop")
wilcox_p <- wilcox.test(per_study$w)$p.value

c(observed = round(w_obs, 2), null_mean = round(mean(w_perm), 3),
  null_sd = round(sd(w_perm), 3),
  studies_positive = sum(per_study$w > 0), of = nrow(per_study),
  spearman = round(cor(wide$zg, wide$zh, method = "spearman"), 2),
  wilcoxon_p = signif(wilcox_p, 2))
```

# Dependence across language models (Supplementary Results 2.5)

For every predictor in the archive paired with GPT-4, on the comparisons where
both are available: the predictor's miss rate, the share of GPT-4's misses it
repeats in the same direction (with a Wilson 95% interval), that pair's exact
corrected-null benchmark for the share, a one-sided binomial P against the
benchmark, and the Pearson correlation between the two predictors'
standardized errors. The binomial tests treat comparisons as independent and
are reported for comparability across predictors; the clustering-robust tests
are in Supplementary Methods 1.7.

```{r crossmodel}
zwide <- all_models %>%
  select(study, outcome.name, reference_condition, condition.name, model, z_recal) %>%
  pivot_wider(names_from = model, values_from = z_recal)
gp <- zwide$`gpt-4`

others <- c("deepseek/deepseek-chat-v3-0324", "openai/gpt-oss-120b", "gpt-3.5-turbo",
            "google/gemma-3-27b-it", "babbage-002", "davinci-002", "human")

cross_template <- map_dfr(others, function(m) {
  zo <- zwide[[m]]
  ok <- !is.na(zo) & !is.na(gp)
  gmr <- mean(abs(gp[ok]) > 1.96)
  omr <- mean(abs(zo[ok]) > 1.96)
  nx  <- null_exact(gmr, omr)
  gmiss <- ok & abs(gp) > 1.96
  k     <- sum((abs(zo) > 1.96 & sign(zo) == sign(gp))[gmiss])
  ci    <- wilson_ci(k, sum(gmiss))
  bench <- nx[["same"]] / gmr
  tibble(predictor = m, miss_rate = omr, n_gpt4_misses = sum(gmiss),
         also_same_side = k / sum(gmiss),
         wilson_lo = ci[["lo"]], wilson_hi = ci[["hi"]],
         null_also_same = bench,
         P_binomial = pbinom(k - 1, sum(gmiss), bench, lower.tail = FALSE),
         error_corr = cor(gp[ok], zo[ok]))
})
cross_template %>%
  mutate(across(c(miss_rate, also_same_side, wilson_lo, wilson_hi,
                  null_also_same, error_corr), ~round(., 2)),
         P_binomial = signif(P_binomial, 2)) %>% as.data.frame()
```

# Secondary archive of megastudies (Supplementary Results 2.6)

The secondary archive contains 606 treatment effects from 15 megastudies, with
expert forecasts available for 376 conditions. The comparison-level file used
here was derived from the authors' `megastudies.RDS` following their scale
conversions, family-specific scale ratios and data corrections, with
predictions recalibrated within each megastudy family by a
leave-one-condition-out procedure (`R/02_build_archive2.R`).

```{r archive2}
a2 <- read_deposited(file.path(derived_dir, "comparisons_archive2.csv"))
a2_total <- c(effects_total = nrow(a2), megastudies = n_distinct(a2$family))
a2_total
a2 <- a2[!is.na(a2$z_recal_gpt) & !is.na(a2$z_recal_expert), ]

gm2   <- abs(a2$z_recal_gpt)    > 1.96
em2   <- abs(a2$z_recal_expert) > 1.96
both2 <- gm2 & em2
same2 <- both2 & (sign(a2$z_recal_gpt) == sign(a2$z_recal_expert))
n2    <- null_exact(mean(gm2), mean(em2))

## every share with its Wilson interval
rbind(gpt_miss              = c(mean(gm2),               wilson_ci(sum(gm2), length(gm2))),
      expert_miss           = c(mean(em2),               wilson_ci(sum(em2), length(em2))),
      expert_also_given_gpt = c(sum(both2) / sum(gm2),   wilson_ci(sum(both2), sum(gm2))),
      same_side_given_both  = c(sum(same2) / sum(both2), wilson_ci(sum(same2), sum(both2))),
      same_dir_given_gpt    = c(sum(same2) / sum(gm2),   wilson_ci(sum(same2), sum(gm2)))
) %>% `colnames<-`(c("share", "lo", "hi")) %>% round(3)

## corrected-null benchmarks for these shares, and the binomial P
bench2 <- n2[["same"]] / mean(gm2)
c(null_expert_also_given_gpt = round(n2[["joint"]] / mean(gm2), 3),
  null_same_side_given_both  = round(n2[["direction"]], 3),
  null_same_dir_given_gpt    = round(bench2, 3),
  P_binomial = signif(pbinom(sum(same2) - 1, sum(gm2), bench2, lower.tail = FALSE), 2))

## archive-wide same-direction joint-miss rate against both benchmarks
arch2_rates <- c(same_side_rate = round(sum(same2) / nrow(a2), 3),
                 null_same_side = round(n2[["same"]], 3),
                 naive_same_side = round(mean(gm2) * mean(em2) / 2, 3))
arch2_rates

## expert-on-model slope, clustered on the megastudy family
sb2 <- slope_band(a2$z_recal_gpt, a2$z_recal_expert, a2$family)
arch2_slope <- c(slope = round(sb2$slope, 2),
                 ci_lo = round(sb2$slope - 1.96 * sb2$slope_se, 2),
                 ci_hi = round(sb2$slope + 1.96 * sb2$slope_se, 2))
arch2_slope
```

With only 15 megastudy-family clusters, the slope's interval should be
interpreted with caution.

# Concentration of same-side joint misses (Supplementary Results 2.7)

```{r concentration}
big <- abs(wide$estimate) >= 0.05          # the authors' five-percentage-point threshold
c(n_larger = sum(big), n_smaller = sum(!big))

mag  <- abs((wide$pg + wide$ph) / 2)       # magnitude of the average raw prediction
gap  <- abs(wide$pg - wide$ph)             # disagreement between the two predictions
cell <- mag >= quantile(mag, .8) & sign(wide$pg) == sign(wide$ph) & gap <= median(gap)

split_stats <- function(flag) {
  k1 <- sum(same[flag]);  n1 <- sum(flag)
  k0 <- sum(same[!flag]); n0 <- sum(!flag)
  c(inside = k1 / n1, wilson_ci(k1, n1),
    outside = k0 / n0, setNames(wilson_ci(k0, n0), c("lo.out", "hi.out")),
    P = prop.test(c(k1, k0), c(n1, n0))$p.value)
}
conc_big  <- split_stats(big)
conc_cell <- split_stats(cell)
rbind(`larger vs smaller effects`   = conc_big,
      `screening cell vs remainder` = conc_cell) %>% signif(3)
c(screening_cell_n = sum(cell))

## direction of the shared errors: the share erring toward zero
toward_zero <- mean(sign(wide$zg[same]) != sign(wide$estimate[same]))
round(toward_zero, 3)
```

This analysis describes where shared misses concentrate; dependence itself is
established by the corrected nulls and the tests in Supplementary Methods
1.4–1.7, because sampling error alone can place both predictions on the same
side of the observed estimate.

# Figures

The three main-text figures are produced by the matplotlib routine below —
the exact code used to render the figures in the manuscript. It reads the
deposited comparison tables, so the numbers in the figures and in the text
cannot drift apart. Rendering the figures requires Python (pandas, numpy,
matplotlib) reachable through **reticulate**; when that is unavailable, the
chunk is skipped and the committed figures in `output/figures/` are shown.

```{r figure-config}
run_py <- isTRUE(params$render_figures) &&
  requireNamespace("reticulate", quietly = TRUE) &&
  tryCatch(reticulate::py_module_available("matplotlib") &&
           reticulate::py_module_available("pandas"),
           error = function(e) FALSE)
ALL_MODELS_CSV <- file.path(derived_dir, "comparisons_all_models.csv")
ARCHIVE2_CSV   <- file.path(derived_dir, "comparisons_archive2.csv")
OUTDIR <- if (dir.exists("../output")) "../output/figures" else "figures"
dir.create(OUTDIR, showWarnings = FALSE, recursive = TRUE)
OUTDIR <- normalizePath(OUTDIR)
if (!run_py) cat("Python figure rendering skipped in this render;",
                 "showing the committed figures.\n")
```

```{python figures, eval=run_py, results='hide'}
#!/usr/bin/env python3
"""
Shared-error reanalysis - main figures.
Fig. 1  GPT-4 vs human crowd (primary archive):   a scatter, b same-way share, c direction split
Fig. 3  GPT-4 vs domain experts (secondary archive): same template
Fig. 2  Each predictor's error vs GPT-4's (primary archive), small multiples

Run:  python3 make_figures.py     (outputs land in ./figures next to this script)
Requires: pandas, numpy, matplotlib
"""
import os, math
import pandas as pd, numpy as np, matplotlib as mpl
mpl.use("Agg")
import matplotlib.pyplot as plt
from matplotlib.lines import Line2D
from matplotlib.patches import Rectangle
from matplotlib.gridspec import GridSpec
from statistics import NormalDist
trapz = getattr(np, 'trapezoid', getattr(np, 'trapz', None))

# ============================ CONFIG ============================
ALL_MODELS_CSV = r.ALL_MODELS_CSV
ARCHIVE2_CSV   = r.ARCHIVE2_CSV
OUTDIR         = r.OUTDIR
BOOT           = 2000        # bootstrap reps for observed-rate / slope CIs
BOOT_NULL      = 400         # bootstrap reps for null CIs (each recomputes the integral)
# ================================================================

os.makedirs(OUTDIR, exist_ok=True)
N = NormalDist()
T = 1.96; CLIP = 12; VIEW = 13.4

mpl.rcParams.update({"font.family":"DejaVu Sans","font.size":9.5,"axes.edgecolor":"#9AA1A9",
    "axes.linewidth":0.8,"xtick.color":"#4A525B","ytick.color":"#4A525B",
    "text.color":"#222222","axes.labelcolor":"#222222"})

RED   = "#A93B2B"   # same-side points / observed bars
BLUE  = "#4C6FA1"   # opposite-side points
LBLUE = "#A9BFD8"   # opposite-direction bar
GRAYP = "#C9CCD1"   # non-joint-miss points
GRAY1 = "#ABABAB"   # corrected-null bar
GRAY2 = "#D7D7D7"   # naive-null bar
INK   = "#222222"
FILL  = "#F6ECEA"   # quadrant shading

clip = lambda v: np.clip(v, -CLIP, CLIP)
_erf = np.vectorize(math.erf)
def Phi(x): return 0.5*(1.0+_erf(np.asarray(x,dtype=float)/1.4142135623730951))

def null_exact(miss_g, miss_h, t=T, ngrid=1201):
    """Corrected null (vectorized): joint-miss rate, same-side rate, direction share
    for two independent predictors scored against the same noisy estimate."""
    if miss_g <= 0 or miss_h <= 0: return {"joint":0.0,"same":0.0,"direction":0.5}
    sig = lambda p: t/N.inv_cdf(1-p/2)
    s2g = max(sig(miss_g)**2-1, 0); s2h = max(sig(miss_h)**2-1, 0)
    u = np.linspace(-9, 9, ngrid); phi = np.exp(-u*u/2)/math.sqrt(2*math.pi)
    def hi(s2): return 1-Phi((t+u)/math.sqrt(s2)) if s2 > 0 else (-u > t).astype(float)
    def lo(s2): return Phi((-t+u)/math.sqrt(s2)) if s2 > 0 else (-u < -t).astype(float)
    gh, gl, hh, hl = hi(s2g), lo(s2g), hi(s2h), lo(s2h)
    joint = trapz((gh+gl)*(hh+hl)*phi, u)
    same  = trapz((gh*hh+gl*hl)*phi, u)
    return {"joint":joint, "same":same, "direction":same/joint}

def cluster_groups(c):
    uq = np.unique(c); return [np.where(c == k)[0] for k in uq]

def slope_ci(x, y, c, B=BOOT):
    np.random.seed(42)
    b = ((x-x.mean())*(y-y.mean())).sum()/((x-x.mean())**2).sum()
    groups = cluster_groups(c); K = len(groups); est = []
    for _ in range(B):
        idx = np.concatenate([groups[j] for j in np.random.randint(0, K, K)])
        xi, yi = x[idx], y[idx]; d = ((xi-xi.mean())**2).sum()
        if d > 0: est.append(((xi-xi.mean())*(yi-yi.mean())).sum()/d)
    return b, *np.percentile(est, [2.5, 97.5])

def boot_stats(x, y, c, B=BOOT, Bn=BOOT_NULL):
    """Cluster bootstrap CIs for: same-way share (of all), direction share (of joint),
    and the two null benchmarks in panel b."""
    np.random.seed(42)
    groups = cluster_groups(c); K = len(groups)
    sw, dr, cn, nn = [], [], [], []
    for i in range(B):
        idx = np.concatenate([groups[j] for j in np.random.randint(0, K, K)])
        xg, yg = x[idx], y[idx]
        gm = np.abs(xg) > T; hm = np.abs(yg) > T; both = gm & hm
        same = both & (np.sign(xg) == np.sign(yg))
        sw.append(same.mean())
        if both.sum() > 0: dr.append(same.sum()/both.sum())
        if i < Bn and gm.mean() > 0 and hm.mean() > 0:
            ne = null_exact(gm.mean(), hm.mean(), ngrid=601)
            cn.append(ne["same"]); nn.append(gm.mean()*hm.mean()*0.5)
    pct = lambda a: (100*np.percentile(a, 2.5), 100*np.percentile(a, 97.5))
    return pct(sw), pct(dr), pct(cn), pct(nn)

# ---------------- the Fig-1 template (used for Fig. 1 and Fig. 3) ----------------
def panel_fig(x, y, c, other_name, unit_word, title_a, outpath):
    n = len(x)
    gm = np.abs(x) > T; hm = np.abs(y) > T; both = gm & hm
    same = both & (np.sign(x) == np.sign(y)); oppo = both & ~same
    over  = same & (x > T); under = same & (x < -T)
    n_joint, n_same, n_oppo = int(both.sum()), int(same.sum()), int(oppo.sum())
    ne = null_exact(gm.mean(), hm.mean())
    obs_b  = 100*same.mean();  corr_b = 100*ne["same"];  naive_b = 100*gm.mean()*hm.mean()*0.5
    obs_c  = 100*n_same/n_joint; oppo_c = 100*n_oppo/n_joint; corr_c = 100*ne["direction"]
    ci_sw, ci_dr, ci_cn, ci_nn = boot_stats(x, y, c)
    ci_op = (100-ci_dr[1], 100-ci_dr[0])

    fig = plt.figure(figsize=(11.7, 6.6), constrained_layout=True)
    gs = GridSpec(2, 2, figure=fig, width_ratios=[1.5, 1], height_ratios=[1, 1])
    axa = fig.add_subplot(gs[:, 0]); axb = fig.add_subplot(gs[0, 1]); axc = fig.add_subplot(gs[1, 1])

    # ---- a: scatter ----
    axa.add_patch(Rectangle((T, T), VIEW-T, VIEW-T, facecolor=FILL, edgecolor="none", zorder=0))
    axa.add_patch(Rectangle((-VIEW, -VIEW), VIEW-T, VIEW-T, facecolor=FILL, edgecolor="none", zorder=0))
    axa.axhline(0, color="#B9BDC2", lw=.8, zorder=1); axa.axvline(0, color="#B9BDC2", lw=.8, zorder=1)
    for s in (T, -T):
        axa.axhline(s, color="#7A828B", lw=.8, ls=(0, (4, 3)), zorder=1)
        axa.axvline(s, color="#7A828B", lw=.8, ls=(0, (4, 3)), zorder=1)
    oth = ~both
    axa.scatter(clip(x[oth]),  clip(y[oth]),  s=5,  c=GRAYP, alpha=.6, lw=0, zorder=2)
    axa.scatter(clip(x[oppo]), clip(y[oppo]), s=11, c=BLUE,  alpha=.9, lw=0, zorder=3)
    axa.scatter(clip(x[same]), clip(y[same]), s=11, c=RED,   alpha=.9, lw=0, zorder=4)
    b, lo, hi = slope_ci(x, y, c); a0 = y.mean()-b*x.mean()
    xx = np.linspace(-CLIP, CLIP, 50)
    axa.fill_between(xx, a0+lo*xx, a0+hi*xx, color="#666666", alpha=.18, lw=0, zorder=5)
    axa.plot(xx, a0+b*xx, color="#1A1A1A", lw=1.6, zorder=6)
    axa.text(-VIEW+0.7,  VIEW-0.8, f"opposite directions\n{n_oppo}",  color=BLUE, fontsize=9, va="top")
    axa.text( VIEW-0.7,  VIEW-0.8, f"both overshoot\n{int(over.sum())}",  color=RED, fontsize=9, va="top", ha="right")
    axa.text(-VIEW+0.7, -VIEW+0.8, f"both undershoot\n{int(under.sum())}", color=RED, fontsize=9, va="bottom")
    axa.text( VIEW-0.7, -VIEW+0.8, f"slope {b:.2f}\n[{lo:.2f}, {hi:.2f}]", color=INK, fontsize=9, va="bottom", ha="right")
    axa.set_xlim(-VIEW, VIEW); axa.set_ylim(-VIEW, VIEW); axa.set_aspect("equal")
    axa.set_xticks([-10, -5, 0, 5, 10]); axa.set_yticks([-10, -5, 0, 5, 10])
    axa.set_xlabel("GPT-4 forecasting error\n(standard errors of the observed effect)", fontsize=9.5)
    axa.set_ylabel(f"{other_name} forecasting error\n(standard errors of the observed effect)", fontsize=9.5)
    axa.set_title(title_a, fontsize=10.5, pad=10)
    for sp in ("top", "right"): axa.spines[sp].set_visible(False)
    axa.text(-0.16, 1.03, "a", transform=axa.transAxes, fontsize=14, fontweight="bold")

    # ---- b: same-way share of all ----
    xs = [0, 1, 2]; vals = [obs_b, corr_b, naive_b]; cols = [RED, GRAY1, GRAY2]
    cis = [ (obs_b, *ci_sw), (corr_b, *ci_cn), (naive_b, *ci_nn) ]
    axb.grid(axis="y", color="#DDDDDD", lw=.7, zorder=0)
    axb.bar(xs, vals, width=.6, color=cols, zorder=2)
    for xpos, (v, l, h) in zip(xs, cis):
        axb.errorbar(xpos, v, yerr=[[max(v-l, 0)], [max(h-v, 0)]], fmt="none",
                     ecolor=INK, elinewidth=1.0, capsize=3, zorder=3)
        axb.text(xpos, h+0.9, f"{v:.1f}%", ha="center", fontsize=9.5, fontweight="bold")
    axb.set_xticks(xs)
    axb.set_xticklabels(["Observed", "Corrected null", "Null ignoring\nshared estimate"], fontsize=8.5)
    top = max(ci_sw[1], obs_b)*1.35
    axb.set_ylim(0, top); axb.set_yticks([0, 10, 20] if top < 30 else [0, 10, 20, 30])
    axb.set_yticklabels([f"{t:.0f}%" if t else "0" for t in axb.get_yticks()])
    axb.set_title(f"Share of all {n:,} {unit_word}\nthat both predictors miss the same way",
                  fontsize=10, loc="left", pad=8)
    for sp in ("top", "right"): axb.spines[sp].set_visible(False)
    axb.text(-0.14, 1.12, "b", transform=axb.transAxes, fontsize=14, fontweight="bold")

    # ---- c: direction split among joint misses ----
    axc.grid(axis="y", color="#DDDDDD", lw=.7, zorder=0)
    axc.bar([0, 1], [obs_c, oppo_c], width=.55, color=[RED, LBLUE], zorder=2)
    for xpos, v, (l, h) in [(0, obs_c, ci_dr), (1, oppo_c, ci_op)]:
        axc.errorbar(xpos, v, yerr=[[max(v-l, 0)], [max(h-v, 0)]], fmt="none",
                     ecolor=INK, elinewidth=1.0, capsize=3, zorder=3)
        axc.text(xpos, h+3.5, f"{v:.0f}%", ha="center", fontsize=9.5, fontweight="bold")
    axc.axhline(corr_c, color=INK, lw=1.2, ls=(0, (5, 2)), zorder=4)
    axc.text(1.42, corr_c+2, f"corrected null, {corr_c:.0f}%", fontsize=8.5, ha="right", color=INK)
    axc.axhline(50, color="#9AA1A9", lw=1.0, ls=(0, (1, 2)), zorder=4)
    axc.text(1.42, 52, "chance, 50%", fontsize=8.5, ha="right", color="#9AA1A9")
    axc.set_xticks([0, 1]); axc.set_xticklabels(["Same\ndirection", "Opposite\ndirections"], fontsize=8.5)
    axc.set_xlim(-.55, 1.55); axc.set_ylim(0, 118); axc.set_yticks([0, 50, 100])
    axc.set_yticklabels(["0", "50%", "100%"])
    axc.set_title(f"Of the {n_joint} {unit_word} both miss,\nthe share missed in the same direction",
                  fontsize=10, loc="left", pad=8)
    for sp in ("top", "right"): axc.spines[sp].set_visible(False)
    axc.text(-0.14, 1.12, "c", transform=axc.transAxes, fontsize=14, fontweight="bold")

    fig.savefig(outpath, dpi=180, bbox_inches="tight")
    plt.close(fig)
    print(f"{os.path.basename(outpath)}: slope {b:.2f} [{lo:.2f},{hi:.2f}] | same-way {obs_b:.1f}% "
          f"(nulls {corr_b:.1f}/{naive_b:.1f}) | direction {obs_c:.0f}% (null {corr_c:.0f})")

# ---------------- Fig 2: small-multiple scatters (unchanged style) ----------------
def scatter_small(ax, x, y, c, title, xlab):
    gm = np.abs(x) > T; em = np.abs(y) > T; both = gm & em
    same = both & (np.sign(x) == np.sign(y)); oppo = both & ~same; oth = ~both
    over = same & (x > T); under = same & (x < -T)
    ax.axhline(0, color="#DfE3E7", lw=.6); ax.axvline(0, color="#DfE3E7", lw=.6)
    for s in (T, -T):
        ax.axhline(s, color="#7A828B", lw=.6, ls=(0, (4, 3))); ax.axvline(s, color="#7A828B", lw=.6, ls=(0, (4, 3)))
    ax.scatter(clip(x[oth]),  clip(y[oth]),  s=5,  c=GRAYP, alpha=.55, lw=0)
    ax.scatter(clip(x[oppo]), clip(y[oppo]), s=7,  c=BLUE,  alpha=.85, lw=0)
    ax.scatter(clip(x[same]), clip(y[same]), s=7,  c=RED,   alpha=.85, lw=0)
    b, lo, hi = slope_ci(x, y, c, B=800); a0 = y.mean()-b*x.mean(); xx = np.linspace(-CLIP, CLIP, 40)
    ax.fill_between(xx, a0+lo*xx, a0+hi*xx, color="#666666", alpha=.16, lw=0)
    ax.plot(xx, a0+b*xx, color="#1A1A1A", lw=1.3)
    ax.text(-VIEW+0.6,  VIEW-0.6, f"opposite\n{int(oppo.sum())}", color=BLUE, fontsize=5.8, va="top", ha="left", linespacing=.9)
    ax.text( VIEW-0.6,  VIEW-0.6, f"both over\n{int(over.sum())}", color=RED, fontsize=5.8, va="top", ha="right", linespacing=.9)
    ax.text(-VIEW+0.6, -VIEW+0.6, f"both under\n{int(under.sum())}", color=RED, fontsize=5.8, va="bottom", ha="left", linespacing=.9)
    ax.text( VIEW-0.6, -VIEW+0.6, f"slope {b:.2f}\n[{lo:.2f}, {hi:.2f}]", color=INK, fontsize=5.8, va="bottom", ha="right", linespacing=.9)
    ax.set_xlim(-VIEW, VIEW); ax.set_ylim(-VIEW, VIEW); ax.set_aspect("equal")
    ax.set_xticks([-10, 0, 10]); ax.set_yticks([-10, 0, 10])
    ax.set_title(title, loc="left", fontweight="bold", fontsize=9.5)
    ax.set_ylabel("forecasting error (SEs)", fontsize=8)
    if xlab: ax.set_xlabel("GPT-4 forecasting error (SEs)", fontsize=8)
    for sp in ("top", "right"): ax.spines[sp].set_visible(False)

# ============================ DATA ============================
am = pd.read_csv(ALL_MODELS_CSV)
key = ["study", "outcome.name", "reference_condition", "condition.name"]
wide = am.pivot_table(index=key, columns="model", values="z_recal", aggfunc="first").reset_index()
g = wide["gpt-4"].values; cl = wide["study"].values

# ---- Fig 1 ----
h = wide["human"].values; ok = ~(np.isnan(g) | np.isnan(h))
panel_fig(g[ok], h[ok], cl[ok], "Human", "comparisons",
          "The two predictors' errors line up, comparison by comparison",
          os.path.join(OUTDIR, "fig1.png"))

# ---- Fig 2 ----
models = [("deepseek/deepseek-chat-v3-0324", "DeepSeek-V3"), ("openai/gpt-oss-120b", "GPT-OSS-120B"),
          ("gpt-3.5-turbo", "GPT-3.5"), ("google/gemma-3-27b-it", "Gemma-3-27B"),
          ("babbage-002", "babbage-002"), ("davinci-002", "davinci-002")]
letters = "abcdef"
fig, axs = plt.subplots(2, 4, figsize=(12.4, 6.9), constrained_layout=True); axs = axs.ravel()
for i, (mc, nm) in enumerate(models):
    zo = wide[mc].values; ok = ~(np.isnan(g) | np.isnan(zo))
    scatter_small(axs[i], g[ok], zo[ok], cl[ok], f"{letters[i]}   {nm}", xlab=(i >= 2))
axs[6].axis("off"); axs[7].axis("off")
LEG = [Line2D([], [], marker='o', ls='', mfc=RED,   mec='none', label='both miss, same side'),
       Line2D([], [], marker='o', ls='', mfc=BLUE,  mec='none', label='both miss, opposite side'),
       Line2D([], [], marker='o', ls='', mfc=GRAYP, mec='none', label='not a joint miss'),
       Line2D([], [], color="#1A1A1A", lw=1.6, label='fitted slope (95% CI)')]
axs[6].legend(handles=LEG, loc="center", frameon=False, fontsize=9,
              title="each point = one of 1,678 comparisons", title_fontsize=9)
fig.savefig(os.path.join(OUTDIR, "fig2.png"), dpi=160, bbox_inches="tight")
plt.close(fig)
print("fig2.png: done")

# ---- Fig 3 ----
a2 = pd.read_csv(ARCHIVE2_CSV).dropna(subset=["z_recal_gpt", "z_recal_expert"])
panel_fig(a2["z_recal_gpt"].values, a2["z_recal_expert"].values, a2["family"].values,
          "Expert", "conditions",
          "The two predictors' errors line up, condition by condition",
          os.path.join(OUTDIR, "fig3.png"))
print("done: fig1.png, fig2.png, fig3.png")
```

```{r show-figures, results='asis', out.width='100%'}
for (f in c("fig1.png", "fig2.png", "fig3.png")) {
  p <- file.path(OUTDIR, f)
  if (file.exists(p)) cat(sprintf("![](%s)\n\n", p))
}
```

# Verification ledger

Every number quoted in the manuscript and its Supplementary Information,
against the value computed in this render. `rule = round` requires the
computed value to equal the quoted one at the quoted precision; `rule = leq`
requires the computed value (a P value bounded by simulation resolution) to
lie at or below the quoted bound; `rule = signif` compares at two significant
digits. Rows that require the raw capsule files are checked only when those
are present. **If any row fails, the render stops here.**

```{r ledger}
rows <- list()
row <- function(section, quantity, quoted, value, digits = 3, rule = "round") {
  rows[[length(rows) + 1]] <<- tibble(section = section, quantity = quantity,
                                      quoted = quoted, computed = as.numeric(value),
                                      digits = digits, rule = rule)
}

## --- Main text ---
row("Main text", "GPT-4 miss rate (recalibrated)", 0.357, mean(miss_g))
row("Main text", "crowd miss rate (recalibrated)", 0.380, mean(miss_h))
row("Main text", "GPT-4 miss rate (unrecalibrated)", 0.463, mean(abs(wide$zg_raw) > 1.96))
row("Main text", "crowd miss rate (unrecalibrated)", 0.498, mean(abs(wide$zh_raw) > 1.96))
row("Main text", "joint misses (n)", 339, sum(joint), 0)
row("Main text", "joint-miss rate", 0.202, mean(joint))
row("Main text", "same-side misses (n)", 305, sum(same), 0)
row("Main text", "same-side rate", 0.182, mean(same))
row("Main text", "direction share | joint miss", 0.90, sum(same) / sum(joint), 2)
row("Main text", "combined recovers | opposite-side miss", 0.82, comb[["recovers_opposite_side"]], 2)
row("Main text", "combined still outside | same-side miss", 0.95, comb[["still_outside_same_side"]], 2)
row("Main text", "corrected null: joint rate", 0.141, ne[["joint"]])
row("Main text", "corrected null: same-side rate", 0.100, ne[["same"]])
row("Main text", "corrected null: direction share", 0.712, ne[["direction"]])
row("Main text", "naive null: same-side rate", 0.068, mean(miss_g) * mean(miss_h) / 2)
row("Main text", "studies with a joint miss", 46, n_distinct(wide$study[joint]), 0)
row("Main text", "same-side rate | larger effects", 0.279, conc_big[["inside"]])
row("Main text", "same-side rate | smaller effects", 0.103, conc_big[["outside"]])
row("Main text", "screening cell size", 75, sum(cell), 0)
row("Main text", "same-side rate inside screening cell", 0.56, conc_cell[["inside"]], 2)
row("Main text", "same-side rate outside screening cell", 0.164, conc_cell[["outside"]])
row("Main text", "share of same-side misses erring toward zero", 0.82, toward_zero, 2)
row("Main text", "crowd-on-GPT-4 error slope", 0.46, sb1$slope, 2)
row("Main text", "expert miss rate (secondary archive)", 0.319, mean(em2))
row("Main text", "GPT-4 miss rate (secondary archive)", 0.303, mean(gm2))
row("Main text", "experts also miss | GPT-4 miss", 0.816, sum(both2) / sum(gm2))
row("Main text", "same side | both miss (secondary)", 0.989, sum(same2) / sum(both2))
row("Main text", "expert-on-GPT-4 error slope", 0.89, sb2$slope, 2)

## --- Supplementary Methods ---
row("SI 1.2", "comparisons in the primary analysis", 1678, nrow(wide), 0)
row("SI 1.2", "studies", 70, n_distinct(wide$study), 0)
row("SI 1.2", "study-outcome combinations", 129,
    wide %>% distinct(study, outcome.name) %>% nrow(), 0)
row("SI 1.3", "pooled recalibration slope, GPT-4", 0.55,
    recal_lines$estimate[recal_lines$model == "gpt-4" & recal_lines$term == "slope"], 2)
row("SI 1.3", "pooled recalibration slope, crowd", 0.62,
    recal_lines$estimate[recal_lines$model == "human" & recal_lines$term == "slope"], 2)
row("SI 1.4", "same-side null, Student t(3) (%)", 9.6, tails["Student t(3)", "same"], 1)
row("SI 1.4", "same-side null, empirical (%)", 9.5, tails["empirical", "same"], 1)
row("SI 1.5", "study-specific null: joint rate", 0.160, ss_rates[["joint"]])
row("SI 1.5", "study-specific null: same-side rate", 0.103, ss_rates[["same_side"]])
row("SI 1.5", "study-specific null: direction share", 0.642, ss_rates[["direction"]])
row("SI 1.6", "bootstrap CI, joint rate, lower", 0.149, boot_ci[1, "joint"])
row("SI 1.6", "bootstrap CI, joint rate, upper", 0.264, boot_ci[2, "joint"])
row("SI 1.6", "bootstrap CI, same-side rate, lower", 0.134, boot_ci[1, "same"])
row("SI 1.6", "bootstrap CI, same-side rate, upper", 0.239, boot_ci[2, "same"])
row("SI 1.6", "bootstrap CI, direction share, lower", 0.841, boot_ci[1, "dir"])
row("SI 1.6", "bootstrap CI, direction share, upper", 0.949, boot_ci[2, "dir"])
row("SI 1.7", "P, joint rate vs corrected null", 1e-4, p_mc[["joint"]], rule = "leq")
row("SI 1.7", "P, same-side rate vs corrected null", 1e-4, p_mc[["same_side"]], rule = "leq")
row("SI 1.7", "P, direction share vs corrected null", 1e-4, p_mc[["direction"]], rule = "leq")
row("SI 1.7", "P, joint rate vs study-specific null", 5e-4, p_ss[["joint"]], rule = "leq")
row("SI 1.7", "P, same-side rate vs study-specific null", 5e-4, p_ss[["same_side"]], rule = "leq")
row("SI 1.7", "P, direction share vs study-specific null", 5e-4, p_ss[["direction"]], rule = "leq")
row("SI 1.7", "bootstrap P, joint rate", 0.010, p_boot[["joint"]], 3)
row("SI 1.7", "bootstrap P, same-side rate", 5e-4, p_boot[["same_side"]], 4)
row("SI 1.7", "bootstrap P, direction share", 2.5e-4, p_boot[["direction"]], rule = "leq")

## --- Supplementary Results ---
row("SI 2.1", "no recalibration: joint rate", 0.276, rob$joint[rob$analysis == "no recalibration"])
row("SI 2.1", "no recalibration: same-side rate", 0.209, rob$same_side[rob$analysis == "no recalibration"])
row("SI 2.1", "error correlation, pp scale, recalibrated", 0.59,
    cor(wide$zg * wide$std.error, wide$zh * wide$std.error), 2)
row("SI 2.1", "error correlation, pp scale, unrecalibrated", 0.39,
    cor(wide$pg - wide$estimate, wide$ph - wide$estimate), 2)
row("SI 2.2", "study-weighted joint rate", 0.302, mean(by_study$joint))
row("SI 2.2", "study-weighted same-side rate", 0.248, mean(by_study$same))
row("SI 2.2", "authors' scheme, 500 draws: joint", 0.23,
    median(draws500[, "joint"]), 2)
row("SI 2.2", "authors' scheme, 500 draws: same-side", 0.20,
    median(draws500[, "same"]), 2)
row("SI 2.2", "s.d. of a single 32-draw run, joint", 0.006, sd(rep32[, "joint"]))
row("SI 2.2", "s.d. of a single 32-draw run, same-side", 0.006, sd(rep32[, "same"]))
row("SI 2.2", "comparisons per draw (median)", 235, median(draws500[, "n"]), 0)
row("SI 2.2", "s.d. across draws, same-side", 0.028, sd(draws500[, "same"]))
row("SI 2.3", "studies with a same-side miss", 43, n_distinct(wide$study[same]), 0)
row("SI 2.3", "slope CI lower (clustered by study)", 0.30, het_slope[["ci_lo"]], 2)
row("SI 2.3", "slope CI upper (clustered by study)", 0.62, het_slope[["ci_hi"]], 2)
row("SI 2.3", "co-error variance between studies (%)", 33.7, varshare, 1)
row("SI 2.3", "studies with >= 5 comparisons", 39, slope_sum[["n_studies"]], 0)
row("SI 2.3", "median per-study slope", 0.58, slope_sum[["median"]], 2)
row("SI 2.3", "per-study slope, 25th percentile", 0.28, slope_sum[["q25"]], 2)
row("SI 2.3", "per-study slope, 75th percentile", 0.97, slope_sum[["q75"]], 2)
row("SI 2.4", "observed mean co-error", 3.62, w_obs, 2)
row("SI 2.4", "permutation s.d.", 0.18, sd(w_perm), 2)
row("SI 2.4", "studies with positive mean co-error", 59, sum(per_study$w > 0), 0)
row("SI 2.4", "Spearman correlation of errors", 0.51,
    cor(wide$zg, wide$zh, method = "spearman"), 2)
row("SI 2.4", "Wilcoxon P (study level)", 5.6e-7, wilcox_p, 2, rule = "signif")
for (i in seq_len(nrow(cross_template))) {
  row("SI 2.5", paste0(cross_template$predictor[i], ": repeats same direction"),
      c(0.78, 0.76, 0.67, 0.64, 0.62, 0.60, 0.51)[i], cross_template$also_same_side[i], 2)
  row("SI 2.5", paste0(cross_template$predictor[i], ": error correlation"),
      c(0.82, 0.79, 0.65, 0.67, 0.57, 0.57, 0.50)[i], cross_template$error_corr[i], 2)
}
row("SI 2.5", "crowd corrected-null benchmark", 0.28,
    cross_template$null_also_same[cross_template$predictor == "human"], 2)
row("SI 2.6", "secondary archive: treatment effects", 606, a2_total[["effects_total"]], 0)
row("SI 2.6", "secondary archive: megastudies", 15, a2_total[["megastudies"]], 0)
row("SI 2.6", "conditions with expert forecasts", 376, nrow(a2), 0)
row("SI 2.6", "null: experts also miss | GPT-4 miss", 0.347, n2[["joint"]] / mean(gm2))
row("SI 2.6", "null: same side | both miss", 0.789, n2[["direction"]])
row("SI 2.6", "same-direction rate, all conditions", 0.245, sum(same2) / nrow(a2))
row("SI 2.6", "null same-direction rate", 0.083, n2[["same"]])
row("SI 2.6", "naive same-direction rate", 0.048, mean(gm2) * mean(em2) / 2)
row("SI 2.6", "expert slope CI lower (clustered)", 0.85, arch2_slope[["ci_lo"]], 2)
row("SI 2.6", "expert slope CI upper (clustered)", 0.93, arch2_slope[["ci_hi"]], 2)
row("SI 2.7", "comparisons with larger effects", 752, sum(big), 0)
row("SI 2.7", "comparisons with smaller effects", 926, sum(!big), 0)

## --- Raw-data counts (checked only when the capsule is present) ---
if (capsule_present) {
  row("SI 1.1", "individual responses", 233601, sum(map_int(rct$data, nrow)), 0)
  row("SI 1.1", "non-missing responses", 229154, sum(map_int(rct$data, ~sum(!is.na(.x$y)))), 0)
  row("SI 1.1", "studies in the raw archive", 71, n_distinct(rct$study), 0)
  row("SI 1.1", "study-outcome cells in the raw archive", 134, nrow(rct), 0)
  row("SI 1.1", "language-model responses", 504840, nrow(llm), 0)
  row("SI 1.1", "responses per model", 72120, nrow(llm) / n_distinct(llm$model), 0)
  row("SI 1.1", "human forecasts", 18317, nrow(fc), 0)
  row("SI 1.1", "forecasters", 2838, n_distinct(fc$PROLIFIC_PID), 0)
  row("SI 1.2", "possible comparisons", 1705,
      rct %>% pmap_int(function(data, ...) {
        k <- n_distinct(data$condition.name)
        as.integer(k * (k - 1) / 2) }) %>% sum(), 0)
}

led <- bind_rows(rows) %>%
  mutate(ok = case_when(
    rule == "round"  ~ abs(round(computed, digits) - quoted) < 1e-9,
    rule == "leq"    ~ computed <= quoted,
    rule == "signif" ~ abs(signif(computed, digits) - quoted) < 1e-12))

led %>%
  transmute(Section = section, Quantity = quantity,
            Quoted   = format(quoted, scientific = FALSE, trim = TRUE, drop0trailing = TRUE),
            Computed = format(signif(computed, 6), scientific = FALSE,
                              trim = TRUE, drop0trailing = TRUE),
            Rule = rule, Match = ifelse(ok, "ok", "FAIL")) %>%
  knitr::kable(align = c("l", "l", "r", "r", "l", "l"))

if (!all(led$ok)) {
  print(filter(led, !ok))
  stop("Verification failed: ", sum(!led$ok),
       " quoted value(s) did not reproduce. See the rows printed above.")
}
cat(sprintf("\nAll %d quoted values reproduce%s.\n", nrow(led),
            if (capsule_present) "" else
            " (raw-data counts not checked: capsule files not present)"))
```

# Session information

```{r session}
sessionInfo()
```
