---
title: "Shared prediction error in Ashokkumar et al. (2026): the full analysis, explained"
author: "Live Leonhardsen Wilhelmsen, Roya Esfandiari & Anton Gollwitzer"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(dplyr); library(tidyr); library(purrr); library(stringr); library(ggplot2)
library(patchwork)   # for composing the multi-panel figures

## Resolved from the project root; place the Code Ocean capsule files (9843791)
## in data/raw/ as described in data/raw/README.md.
library(here)
data_dir <- here::here("data", "raw")
```

# What this document is

This is the complete, self-contained analysis behind our Matters Arising on
Ashokkumar, Hewitt, Ghezae & Willer, *Large language models can predict the
results of social science experiments* (Nature, 2026). Every number quoted in
the commentary and its Supplementary Information is computed here, in order,
with an explanation of what each step does and why we do it that way.

**The question in one sentence.** The authors show that GPT-4 and a crowd of
human forecasters each predict experimental results about equally well, and
suggest the two can be combined because they carry partly independent
information. We ask about the part they did not analyse: the *errors*. When
the model is wrong and the crowd is wrong, are they wrong on the same
experiments, in the same direction? If yes, averaging them cannot fix those
cases (same-direction errors do not cancel), and their agreement is not
evidence that a prediction is right.

# The data

Three files from the authors' capsule are used. Nothing here modifies them.

```{r load-data}
rct <- readRDS(file.path(data_dir, "rct_responses.RDS"))
llm <- readRDS(file.path(data_dir, "llm_responses.RDS"))
fc  <- readRDS(file.path(data_dir, "forecasting_responses.RDS"))

c(rct_responses   = sum(map_int(rct$data, nrow)),   # individual survey answers
  llm_rows        = nrow(llm),                       # one row per model call (7 models)
  forecast_rows   = nrow(fc),                        # one row per forecaster x condition
  forecasters     = n_distinct(fc$PROLIFIC_PID))

## descriptive counts referenced in the Supplementary Information
c(answers_nonmissing = sum(map_int(rct$data, ~sum(!is.na(.x$y)))),
  studies_raw        = n_distinct(rct$study),
  study_outcome_cells_raw = nrow(rct),
  conditions_gpt4    = llm %>% filter(model == "gpt-4") %>%
                         distinct(study, condition.name) %>% nrow(),
  possible_pairs     = rct %>% pmap_int(function(data, ...) {
                         k <- n_distinct(data$condition.name)
                         as.integer(k * (k - 1) / 2) }) %>% sum(),
  prompt_cells_240   = llm %>% count(model, study, outcome.name, condition.name) %>%
                         filter(model == "gpt-4", n == 240) %>% nrow())
```

*What you are looking at:* `rct_responses` holds the real answers of the
~120,000 experiment participants — this is the ground truth the predictors are
trying to guess. `llm_responses` holds every single model call (about 120
prompts per experimental condition, each prompt a demographic profile, each
recorded value already the average of five samples). `forecasting_responses`
holds the human predictions — roughly 30 forecasters per question.

# Step 1 — the observed effects

**What we do.** For every experiment, every outcome and every *pair* of
conditions, we compute the observed treatment effect: the difference in mean
response between the two conditions, after rescaling every outcome to run from
0 to 1 so that different answer scales are comparable. We also keep the
effect's **standard error** — the experiment's own measure of how precisely it
pinned that difference down.

**Why a regression?** Running `lm(y ~ condition)` inside each study–outcome is
just a convenient way to get every condition-versus-reference difference *and*
its standard error in one step; the coefficient literally equals the difference
in means. This is exactly what the authors' own pipeline does, so our observed
effects are theirs.

```{r observed-effects, cache=TRUE}
## One technical note: we recode condition names to plain ASCII codes before the
## regression and map back afterwards. Some condition names contain typographic
## quotes, and on computers with a non-UTF-8 locale R silently rewrites such
## characters inside coefficient names, which would break the merge with the
## prediction files. The recode makes the analysis give identical results on
## any machine.
rct_effects <- rct %>%
  pmap_dfr(function(study, outcome.name, outcome.min, outcome.max, data, ...) {
    d <- data %>% ungroup() %>%
      mutate(y = (as.numeric(y) - outcome.min) / (outcome.max - outcome.min))
    conds <- unique(d$condition.name)
    code  <- setNames(paste0("k", seq_along(conds)), conds)   # safe ASCII level codes
    d$cc  <- code[d$condition.name]
    map_dfr(conds, function(ref) {
      fit <- lm(y ~ cc, data = d %>% mutate(cc = forcats::fct_relevel(cc, code[[ref]])))
      broom::tidy(fit) %>%
        transmute(cc = str_match(term, "^cc(k\\d+)$")[, 2], estimate, std.error) %>%
        filter(!is.na(cc)) %>%
        mutate(condition.name = names(code)[match(cc, code)],
               study = study, outcome.name = outcome.name,
               reference_condition = ref, .before = 1) %>%
        select(-cc)
    })
  })
```

# Step 2 — the predictions

**The model.** For GPT-4 (and later the six other models), the predicted mean
of a condition is the **weighted** average of all its prompt-level responses.
The weights matter: the authors over-sampled some demographic groups so they
could study subgroups, and the weights undo that over-sampling so the ensemble
represents the US population — we use their weights exactly. Reversed answer
scales are un-flipped first, again following their code.

**The crowd.** Each forecaster predicted the average answer of a condition on
a 0–100 slider; the crowd prediction is the plain mean of those forecasts,
divided by 100.

**A predicted effect** is then simply: predicted mean of condition A minus
predicted mean of condition B — the same subtraction the experiment performs
with real people.

```{r predictions, cache=TRUE}
llm_means <- llm %>%
  mutate(expectation = ifelse(scale_flip, outcome_scale_max - (expectation - outcome_scale_min),
                              expectation),
         y = (expectation - outcome_scale_min) / (outcome_scale_max - outcome_scale_min)) %>%
  filter(!is.na(y)) %>%
  group_by(model, study, outcome.name, condition.name) %>%
  summarise(pred_mean = weighted.mean(y, w = weight), .groups = "drop")

fc_means <- fc %>%
  mutate(y = value / 100) %>%
  semi_join(distinct(llm, study, outcome.name), by = c("study", "outcome.name")) %>%
  group_by(study, outcome.name, condition.name) %>%
  summarise(pred_mean = mean(y), .groups = "drop") %>%
  mutate(model = "human")

cond_means <- bind_rows(llm_means, fc_means)

pred_contrasts <- cond_means %>%
  group_by(model, study, outcome.name) %>%
  group_modify(function(d, k) {
    crossing(d %>% select(condition.name, pred_mean),
             d %>% select(reference_condition = condition.name, ref_mean = pred_mean)) %>%
      filter(condition.name != reference_condition) %>%
      transmute(condition.name, reference_condition, prediction = pred_mean - ref_mean)
  }) %>%
  ungroup()
```

# Step 3 — one row per comparison, and the 'combined' predictor

Comparing A to B and B to A is the same comparison with the sign flipped, so we
keep each **unordered pair** once. The authors' proposed combination of model
and crowd is the simple average of the two predictions; we build it the same
way so we can test it later.

```{r comparisons, cache=TRUE}
## Deduplication note: comparing strings with '<' depends on the computer's
## locale — in a plain C locale R even returns NA for names with typographic
## quotes, silently dropping those comparisons. We therefore order names with a
## fixed ICU collator (English), which is identical on every machine and
## matches the orientation of the deposited comparison-level dataset.
en <- stringi::stri_opts_collator(locale = "en_US")
comparisons <- rct_effects %>%
  inner_join(pred_contrasts, by = c("study", "outcome.name", "condition.name", "reference_condition")) %>%
  filter(stringi::stri_cmp_lt(reference_condition, condition.name, opts_collator = en))

combined <- comparisons %>%
  filter(model %in% c("gpt-4", "human")) %>%
  select(study, outcome.name, condition.name, reference_condition, estimate, std.error, model, prediction) %>%
  pivot_wider(names_from = model, values_from = prediction) %>%
  filter(!is.na(`gpt-4`), !is.na(human)) %>%
  mutate(model = "combined", prediction = (`gpt-4` + human) / 2) %>%
  select(-`gpt-4`, -human)

comparisons <- bind_rows(comparisons, combined)
```

# Step 4 — how wrong is a prediction? The z-score

**The idea.** A prediction that is 5 points off means nothing in a small noisy
study and a lot in a big precise one. So we measure each prediction's error in
units of the observed effect's own standard error:

$$z = \frac{\text{prediction} - \text{observed effect}}{\text{standard error of the observed effect}}$$

If the prediction were perfect and only sampling noise separated it from the
estimate, z would behave like a standard normal draw: beyond ±1.96 only 5% of
the time. So **|z| > 1.96 means the prediction sits outside the experiment's
own 95% confidence interval — the experiment itself is telling us the
prediction is wrong.** We call that a *miss*.

# Step 5 — recalibration (giving the predictors their best shot)

**Why.** The authors themselves report that predictions overshoot true effects
by roughly a factor of two, and that a linear rescaling fixes it. If we skipped
this, a critic could say all our shared misses are just that one shared,
already-known bias. So we apply *their* remedy before measuring anything.

**How, without cheating.** For each study we fit the rescaling line
`lm(estimate ~ prediction)` on the *other 69* studies only, then apply it to
the held-out study — so no study helps correct itself, and every predictor gets
the same treatment.

**Where this procedure comes from.** To be precise about provenance: this
leave-one-study-out rescaling is the authors' method for their *pilot-testing*
use-case, not part of the main correlation analysis they report in the paper.
We borrow it because it is their own published remedy for the overshooting they
document, and applying it is the version of the analysis most favourable to the
predictors. Importantly, the shared error is **not** created by it — it is
already present in the raw, unrecalibrated predictions (see the robustness
section: joint misses 27.6%, same-side 20.9%, and an error correlation of 0.39
before recalibration versus 0.59 after). Recalibration sharpens the picture; it
does not manufacture it.

```{r recalibration, cache=TRUE}
comparisons <- comparisons %>%
  mutate(z_raw = (prediction - estimate) / std.error) %>%
  group_by(model) %>%
  group_modify(function(d, k) {
    d$pred_recal <- NA_real_
    for (s in unique(d$study)) {
      fit <- lm(estimate ~ prediction, data = d %>% filter(study != s))
      d$pred_recal[d$study == s] <- predict(fit, newdata = d %>% filter(study == s))
    }
    d
  }) %>%
  ungroup() %>%
  mutate(z_recal = (pred_recal - estimate) / std.error)

## the pooled recalibration lines (the leave-one-study-out fits vary only
## slightly around these) — intercepts are essentially zero, slopes about 0.5,
## i.e. both predictors overstate effects by roughly a factor of two.
comparisons %>%
  filter(model %in% c("gpt-4", "human")) %>%
  group_by(model) %>%
  group_modify(~ broom::tidy(lm(estimate ~ prediction, data = .x))) %>%
  ungroup() %>%
  transmute(model, term, estimate = round(estimate, 4))

wide <- comparisons %>%
  filter(model %in% c("gpt-4", "human", "combined")) %>%
  select(study, outcome.name, reference_condition, condition.name,
         estimate, std.error, prediction, model, z_raw, z_recal) %>%
  pivot_wider(names_from = model, values_from = c(prediction, z_raw, z_recal)) %>%
  filter(!is.na(`z_recal_gpt-4`), !is.na(z_recal_human)) %>%
  rename(zg = `z_recal_gpt-4`, zh = z_recal_human, zc = z_recal_combined,
         pg = `prediction_gpt-4`, ph = prediction_human)

nrow(wide)   # 1,678 — all available comparisons
```

# The core results

A note on what the 1,678 are: they are **all available comparisons** — every
pair of conditions within a study and outcome for which both predictors made a
prediction. They are deliberately not a random or independent sample. Pairs
within a study share participants and often share a condition, so they overlap;
this is why every confidence interval below is computed by resampling *studies*
rather than comparisons, and why we also report study-weighted rates and the
authors' own sampling scheme as checks.

```{r core}
miss_g <- abs(wide$zg) > 1.96
miss_h <- abs(wide$zh) > 1.96
joint  <- miss_g & miss_h
same   <- joint & (sign(wide$zg) == sign(wide$zh))

tibble(
  `GPT-4 miss rate`  = mean(miss_g),
  `Crowd miss rate`  = mean(miss_h),
  `Joint misses`     = sum(joint),
  `Joint rate`       = mean(joint),
  `Same-side misses` = sum(same),
  `Same-side rate`   = mean(same),
  `Direction | joint`= sum(same) / sum(joint)
) %>% mutate(across(where(is.numeric), ~round(., 3))) %>% as.data.frame()

## the analysis set behind the 1,678, and a rank-based coupling measure
c(studies = n_distinct(wide$study),
  study_outcome_cells = wide %>% distinct(study, outcome.name) %>% nrow(),
  spearman_zg_zh = round(cor(wide$zg, wide$zh, method = "spearman"), 2))
```

*Reading this:* even after the correction, each predictor misses over a third
of the time (35.7% and 38.0%). They miss **together** on 339 comparisons
(20.2%), and on 305 of those (18.2% of everything) they miss **on the same
side** — both too high, or both too low, relative to what the experiment found.
Of the comparisons both miss, 90% are same-side.

```{r scatter, fig.width=6, fig.height=5}
wide %>%
  mutate(class = case_when(same ~ "Both miss, same side",
                           joint ~ "Both miss, opposite sides",
                           TRUE ~ "Other")) %>%
  ggplot(aes(pmax(pmin(zg, 12), -12), pmax(pmin(zh, 12), -12), colour = class)) +
  geom_point(size = .8, alpha = .6) +
  geom_hline(yintercept = c(-1.96, 1.96), linetype = 3) +
  geom_vline(xintercept = c(-1.96, 1.96), linetype = 3) +
  scale_colour_manual(values = c("Both miss, same side" = "#c0392b",
                                 "Both miss, opposite sides" = "#1f6eb2",
                                 "Other" = "grey75")) +
  labs(x = "GPT-4 error (standard errors of the observed effect)",
       y = "Human crowd error", colour = NULL) +
  theme_minimal()
```

# Are the errors dependent? A test with no assumptions

**The idea.** Multiply the two errors together, comparison by comparison:
$w = z_g \times z_h$. When the two err on the same side, w is positive; on
opposite sides, negative. If the two predictors were unrelated, positives and
negatives would balance and the average w would be zero. To get the "unrelated"
benchmark without assuming anything, we simply **shuffle the pairing**: match
GPT-4's error on one comparison with the crowd's error on a random other
comparison, 2,000 times. Shuffling destroys any true dependence while keeping
each predictor's own error distribution exactly as it is.

```{r permutation}
set.seed(1)
w_obs <- mean(wide$zg * wide$zh)
w_perm <- replicate(2000, mean(wide$zg * sample(wide$zh)))
c(observed = round(w_obs, 2), null_mean = round(mean(w_perm), 3), null_sd = round(sd(w_perm), 3))

per_study <- wide %>% group_by(study) %>% summarise(w = mean(zg * zh))
c(studies_positive = sum(per_study$w > 0), of = nrow(per_study),
  wilcoxon_p = wilcox.test(per_study$w)$p.value)
```

*Reading this:* the observed mean co-error is 3.6 where shuffling gives 0 —
the errors are strongly dependent. And it is not one weird study: the study-level
mean is positive in 59 of the 70. The Wilcoxon test makes the point while
treating the study, not the comparison, as the unit — the conservative choice,
since comparisons inside a study share participants.

# The corrected null: what would "unrelated" really look like?

**The subtlety.** Both predictors are graded against the *same* estimated
effect, and that estimate contains sampling noise of its own. When an
experiment's estimate happens by chance to land unusually high, *every*
prediction ends up "too low" relative to it — on the same side — through no
fault of the predictors. So even completely unrelated predictions would show
some same-side clustering. Naive benchmarks (50% direction; the product of the
two miss rates) ignore this and would flatter us.

**The fix.** Write each standardized error as $z_i = a_i - u$, where $a_i$ is
the predictor's own deviation from the *true* effect and $u$ is the estimate's
standardized sampling noise — the same $u$ for both predictors, with variance
exactly 1 by construction. We simulate unrelated predictors that share only
$u$, calibrating each one's spread so its overall miss rate matches the real
predictor's. What this null produces is the benchmark; whatever the data show
*beyond* it is genuine dependence between the predictions.

```{r corrected-null, cache=TRUE}
set.seed(2)
N <- 2e6
u <- rnorm(N); bg <- rnorm(N); bh <- rnorm(N)
calib <- function(b, target)
  uniroot(function(s) mean(abs(s * b - u) > 1.96) - target, c(0.05, 20))$root
sg <- calib(bg, mean(miss_g)); sh <- calib(bh, mean(miss_h))
zgs <- sg * bg - u; zhs <- sh * bh - u
joint_s <- abs(zgs) > 1.96 & abs(zhs) > 1.96
same_s  <- joint_s & sign(zgs) == sign(zhs)
c(null_joint = round(mean(joint_s), 3),
  null_same  = round(mean(same_s), 3),
  null_direction = round(mean(sign(zgs[joint_s]) == sign(zhs[joint_s])), 3))

## sanity check: theory says E[z_g z_h] = Var(u) = 1 exactly under this null
round(mean(zgs * zhs), 3)
```

*Reading this:* under the corrected null, unrelated predictors would already
produce **71%** same-side agreement among joint misses (not 50%) and a **10%**
same-side miss rate (not 6.8%). The observed values — 90% and 18.2% — exceed
both. That is the honest comparison, and it is the one the commentary reports.

**An exact version.** Under this null the pair of errors is exactly bivariate
normal, so the benchmarks are one-dimensional integrals and need no simulation
at all. The function below computes them exactly for any pair of miss rates and
any threshold; we use it for every benchmark in the Supplementary Information,
so those numbers carry no simulation noise. It reproduces the simulated values
above to three decimals.

```{r exact-null}
null_exact <- function(miss_g, miss_h, t = 1.96) {
  sig <- function(p) t / qnorm(1 - p / 2)            # total sd that yields this miss rate
  s2g <- max(sig(miss_g)^2 - 1, 0); s2h <- max(sig(miss_h)^2 - 1, 0)
  tail_hi <- function(s2, u) if (s2 > 0) pnorm((t + u) / sqrt(s2), lower.tail = FALSE) else as.numeric(-u >  t)
  tail_lo <- function(s2, u) if (s2 > 0) pnorm((-t + u) / sqrt(s2))                    else as.numeric(-u < -t)
  f <- function(u, kind) sapply(u, function(uu) {
    gh <- tail_hi(s2g, uu); gl <- tail_lo(s2g, uu)
    hh <- tail_hi(s2h, uu); hl <- tail_lo(s2h, uu)
    v <- if (kind == "joint") (gh + gl) * (hh + hl) else gh * hh + gl * hl
    v * dnorm(uu)
  })
  joint <- integrate(f, -9, 9, kind = "joint", rel.tol = 1e-9, subdivisions = 500L)$value
  same  <- integrate(f, -9, 9, kind = "same",  rel.tol = 1e-9, subdivisions = 500L)$value
  c(joint = joint, same = same, direction = same / joint)
}
round(100 * null_exact(mean(miss_g), mean(miss_h)), 2)   # matches the simulation above
```

**Does the null's normality matter?** The predictors' real errors have heavier
tails than a normal distribution, so we recompute the two decisive benchmarks
with heavier-tailed error shapes — a Student t with 3 degrees of freedom, and
the empirical shape of the errors themselves. Both are again computed exactly,
by integrating over the shared term. Heavier tails make the benchmarks slightly
*lower*, so the normal case quoted in the commentary is the most conservative.

```{r tail-variants, cache=TRUE}
null_exact_tail <- function(miss_g, miss_h, t = 1.96, qtail, ptail) {
  ## ptail(q): upper-tail prob of the standardized error shape; calibrate scale s
  ## so that P(|s*b - u| > t) matches each predictor's observed miss rate.
  miss_of <- function(s) integrate(function(u) sapply(u, function(uu)
      (ptail((t + uu) / s) + ptail((t - uu) / s)) * dnorm(uu)), -9, 9,
      subdivisions = 500L, stop.on.error = FALSE)$value
  cal <- function(p) uniroot(function(s) miss_of(s) - p, c(0.05, 30))$root
  sg <- cal(miss_g); sh <- cal(miss_h)
  f <- function(u, kind) sapply(u, function(uu) {
    gh <- ptail((t + uu) / sg); gl <- ptail((t - uu) / sg)
    hh <- ptail((t + uu) / sh); hl <- ptail((t - uu) / sh)
    v <- if (kind == "joint") (gh + gl) * (hh + hl) else gh * hh + gl * hl
    v * dnorm(uu)
  })
  joint <- integrate(f, -9, 9, kind = "joint", subdivisions = 500L, stop.on.error = FALSE)$value
  same  <- integrate(f, -9, 9, kind = "same",  subdivisions = 500L, stop.on.error = FALSE)$value
  c(same = 100 * same, direction = 100 * same / joint)
}
t3p <- function(q) pt(q * sqrt(3), df = 3, lower.tail = FALSE)   # t(3) scaled to unit variance
eg  <- (wide$zg - mean(wide$zg)) / sd(wide$zg)                   # empirical shapes
eh  <- (wide$zh - mean(wide$zh)) / sd(wide$zh)
emp <- function(z) function(q) sapply(q, function(qq) mean(z > qq))
rbind(
  `normal`        = 100 * null_exact(mean(miss_g), mean(miss_h))[c("same", "direction")] * c(1, 1),
  `Student t(3)`  = null_exact_tail(mean(miss_g), mean(miss_h), qtail = NULL, ptail = t3p),
  `empirical`     = null_exact_tail(mean(miss_g), mean(miss_h), qtail = NULL,
                                    ptail = function(q) (emp(eg)(q) + emp(eh)(q)) / 2)
) %>% round(1)
```

# A fairer benchmark: letting each study set its own difficulty

**The worry.** The corrected null above gives each predictor a single miss rate
for the whole archive — in effect assuming every study is equally hard to
predict. But studies plainly differ: some test effects almost anyone would call
correctly, others are genuinely obscure. If a study is simply *hard*, both
predictors will miss it more often **even if their errors have nothing to do
with each other**, and that extra overlap would show up in our numbers as
"shared error" when it is really just "hard study".

**The fix.** Run exactly the same null, but calibrate the two spreads *within
each study* to that study's own miss rates, then pool across studies. This
builds study difficulty into the benchmark itself, so whatever the data show
above it cannot be explained by some studies being harder than others.

```{r study-specific-null, cache=TRUE}
set.seed(21)
POOL   <- 1e6
pool_b <- rnorm(POOL); pool_u <- rnorm(POOL)
rate_for <- function(s) mean(abs(s * pool_b - pool_u) > 1.96)

## calibrate the spread s so the simulated miss rate matches a target.
## Guards: a perfect predictor still "misses" 5% of the time (that is the floor),
## and we cap the spread at 40 for numerical safety.
calib_s <- function(target) {
  if (target <= rate_for(0))  return(0)
  if (target >= rate_for(40)) return(40)
  uniroot(function(s) rate_for(s) - target, c(0, 40), tol = 1e-4)$root
}

REP <- 4000                      # simulated replicates per real comparison
ss_null <- map_dfr(unique(wide$study), function(k) {
  m  <- wide$study == k
  nk <- sum(m)
  s1 <- calib_s(mean(miss_g[m]))
  s2 <- calib_s(mean(miss_h[m]))
  uu <- rnorm(nk * REP); b1 <- rnorm(nk * REP); b2 <- rnorm(nk * REP)
  z1 <- s1 * b1 - uu; z2 <- s2 * b2 - uu
  j  <- abs(z1) > 1.96 & abs(z2) > 1.96
  tibble(n = length(z1), joint = sum(j), same = sum(j & sign(z1) == sign(z2)))
})

rbind(
  observed             = c(mean(joint), mean(same), sum(same) / sum(joint)),
  `global null`        = c(mean(joint_s), mean(same_s),
                           mean(sign(zgs[joint_s]) == sign(zhs[joint_s]))),
  `study-specific null`= c(sum(ss_null$joint) / sum(ss_null$n),
                           sum(ss_null$same)  / sum(ss_null$n),
                           sum(ss_null$same)  / sum(ss_null$joint))
) %>% `colnames<-`(c("joint", "same_side", "direction")) %>% round(3)
```

*Reading this:* the joint-miss benchmark rises from about 14% to about 16%, so
study difficulty does explain part of the overlap in **which** comparisons get
missed — roughly a third of the excess. But the same-side benchmark barely
moves (10.0% to about 10.3%), and the direction benchmark actually *falls*
(71% to about 64%), so the observed 90% clears it by more than before.

That is exactly the pattern the logic predicts. A hard study makes both
predictors miss the same *items*, but being hard gives them no reason to miss
on the same *side*. So the directional result — the one that decides whether
averaging can help — survives this objection untouched, while the joint-miss
excess, which we already treated as the weaker of the two, narrows.

One technical caveat: in a few studies a predictor misses so rarely that no
spread can reproduce its rate (a perfect predictor still lands outside the
interval 5% of the time by chance). We set the spread to zero there, which
makes the two simulated predictors perfectly aligned and so *inflates* the
null — the conservative direction. This affects 4 studies (9 comparisons, with
zero observed joint misses between them), so it cannot be driving the result.

# How certain are we? Uncertainty that respects the studies

Comparisons inside a study are not independent — they share participants and
conditions. So instead of pretending we have 1,678 independent observations,
we bootstrap **studies**: resample the 70 studies with replacement, recompute
each rate, repeat 4,000 times, and take the middle 95%.

```{r bootstrap, cache=TRUE}
set.seed(3)
studies <- unique(wide$study)
boot <- replicate(4000, {
  s <- sample(studies, replace = TRUE)
  d <- map_dfr(s, ~ wide[wide$study == .x, ])
  jg <- abs(d$zg) > 1.96 & abs(d$zh) > 1.96
  sm <- jg & sign(d$zg) == sign(d$zh)
  c(joint = mean(jg), same = mean(sm), dir = sum(sm) / sum(jg))
})
apply(boot, 1, quantile, c(.025, .975)) %>% round(3)
```

*Reading this:* the same-side rate's interval (about 13–24%) clears the 10%
null, and the direction interval (about 84–95%) clears 71% comfortably. The
joint-miss interval brushes its 14% null — which is why the commentary calls
that excess modest and leans on direction.

# Is one big study driving the result? Study-weighted rates

Studies contribute very unequal numbers of comparisons — from 1 to 198 — so the
headline rates are dominated by the condition-rich studies. The check is to
recompute them giving each of the 70 studies equal weight.

```{r study-weighted}
by_study <- wide %>%
  group_by(study) %>%
  summarise(n     = n(),
            joint = mean(abs(zg) > 1.96 & abs(zh) > 1.96),
            same  = mean(abs(zg) > 1.96 & abs(zh) > 1.96 & sign(zg) == sign(zh)),
            .groups = "drop")

c(joint_comparison_weighted = mean(joint),
  joint_study_weighted      = mean(by_study$joint),
  same_comparison_weighted  = mean(same),
  same_study_weighted       = mean(by_study$same)) %>% round(3)

range(by_study$n)      # comparisons contributed by the smallest and largest study
```

*Reading this:* the study-weighted rates are **higher** (about 30% and 25%)
than the comparison-weighted ones we report (20.2% and 18.2%). So the large,
condition-rich studies are pulling our headline figures *down*, not up: the
numbers in the commentary are the conservative choice, and giving every study
an equal say would strengthen the result rather than weaken it.

# Does averaging actually fail? Their own combined predictor

No assumptions needed here either: the authors propose averaging model and
crowd, we built that average above and recalibrated it identically — so we can
just look at how it does on the comparisons in question.

```{r combined}
tibble(
  `same-side joint misses: combined also outside CI` =
    mean(abs(wide$zc[same]) > 1.96),
  `opposite-side joint misses: combined outside CI` =
    mean(abs(wide$zc[joint & !same]) > 1.96)
) %>% round(3)
```

*Reading this:* where the two miss on opposite sides, averaging works —
the combined prediction recovers 82% of the time. Where they miss on the same
side, it stays wrong 95% of the time. Same-side error is precisely the failure
mode averaging cannot repair, in their own combination.

# Can a different model, or an expert, be the independent check?

```{r crossmodel}
zwide <- comparisons %>%
  select(study, outcome.name, reference_condition, condition.name, model, z_recal) %>%
  pivot_wider(names_from = model, values_from = z_recal)

others <- c("deepseek/deepseek-chat-v3-0324", "openai/gpt-oss-120b", "gpt-3.5-turbo",
            "google/gemma-3-27b-it", "babbage-002", "davinci-002", "human")
gp <- zwide$`gpt-4`
map_dfr(others, function(m) {
  zo <- zwide[[m]]
  ok <- !is.na(zo) & !is.na(gp)
  gmr <- mean(abs(gp[ok]) > 1.96)                     # GPT-4's miss rate on this overlap
  omr <- mean(abs(zo[ok]) > 1.96)
  ne  <- null_exact(gmr, omr)                         # exact corrected null for this pair
  gmiss <- ok & abs(gp) > 1.96
  tibble(predictor = m,
         miss_rate = omr,
         also_same_side = mean((abs(zo) > 1.96 & sign(zo) == sign(gp))[gmiss]),
         null_also_same = ne[["same"]] / gmr,
         error_corr = cor(gp[ok], zo[ok]))
}) %>% mutate(across(where(is.numeric), ~round(., 2)))
```

*Reading this:* every language model — including ones from unrelated companies
and older generations — repeats GPT-4's same-side misses far more often (60–78%)
than an unrelated predictor would (about 29–31% under the corrected null,
which is high only because of the shared yardstick). The human crowd, at 51%,
is the *least* entangled predictor available. There is no independent second
opinion on this list.

## And domain experts? The second archive

The authors' secondary archive holds 15 megastudies whose original teams
collected forecasts from domain experts. The comparison-level file used here
(`comparisons_archive2.csv`) was built from the authors' `megastudies.RDS` by
our archive-2 script, following their own scale conversions and dataset
corrections, with recalibration done within megastudy family. The question is
the same: do experts miss where the model misses?

```{r archive2}
a2 <- read.csv(file.path(data_dir, "comparisons_archive2.csv"), check.names = FALSE)
c(effects_total = nrow(a2), megastudies = n_distinct(a2$family))   # the full second archive
a2 <- a2[!is.na(a2$z_recal_gpt) & !is.na(a2$z_recal_expert), ]   # conditions with expert forecasts

gm2   <- abs(a2$z_recal_gpt)    > 1.96
em2   <- abs(a2$z_recal_expert) > 1.96
both2 <- gm2 & em2
same2 <- both2 & (sign(a2$z_recal_gpt) == sign(a2$z_recal_expert))

c(conditions            = nrow(a2),
  gpt_miss              = round(mean(gm2), 3),
  expert_miss           = round(mean(em2), 3),
  expert_also_given_gpt = round(sum(both2) / sum(gm2), 3),
  same_side_given_both  = round(sum(same2) / sum(both2), 3),
  slope_expert_on_gpt   = round(coef(lm(a2$z_recal_expert ~ a2$z_recal_gpt))[2], 2))

## the corrected-null benchmarks for these two shares:
n2 <- null_exact(mean(gm2), mean(em2))
c(null_expert_also_given_gpt = round(n2[["joint"]] / mean(gm2), 3),
  null_same_side_given_both  = round(n2[["direction"]], 3))
```

*Reading this:* on the 376 conditions with expert forecasts, experts miss about
as often as the model (31.9% versus 30.3%); of the conditions GPT-4 misses,
experts also miss 82%, and 99% of those on the same side. Expertise does not
supply the independent second opinion either.

# Where the shared misses fall

**Why this section matters.** A user screening interventions or flagging
effects for replication acts where the predictions look big and agreed. If
shared misses were scattered at random they would be an accuracy tax; if they
concentrate exactly where a user would act, they are a selection hazard.

```{r concentration}
big <- abs(wide$estimate) >= 0.05          # the authors' 5-percentage-point threshold
tibble(`same-side rate | larger effects` = mean(same[big]),
       `same-side rate | smaller effects` = mean(same[!big]),
       n_larger = sum(big), n_smaller = sum(!big)) %>% round(3)

mag  <- abs((wide$pg + wide$ph) / 2)       # size of the (uncorrected) average prediction
gap  <- abs(wide$pg - wide$ph)             # how much the two disagree
cell <- mag >= quantile(mag, .8) & sign(wide$pg) == sign(wide$ph) & gap <= median(gap)
tibble(cell_n = sum(cell),
       `same-side rate inside cell` = mean(same[cell]),
       `same-side rate outside` = mean(same[!cell])) %>% round(3)

## and the direction of the shared errors:
mean(sign(wide$zg[same]) != sign(wide$estimate[same]))   # share erring toward zero

## how widely the shared misses are spread across the 70 studies
c(studies_with_joint_miss = n_distinct(wide$study[joint]),
  studies_with_same_side  = n_distinct(wide$study[same]))
```

*Reading this:* the misses are ~3x as common among the larger observed effects
(27.9% vs 10.3%), and in the "looks most trustworthy" cell — big predicted
effect, both predictors agreeing closely in sign and size — 56% of comparisons
are same-side misses. And 82% of all same-side misses have both predictors
erring toward zero: shared conservatism about large effects.

# Robustness: none of this depends on our choices

Three alternative analyses, same conclusion. (1) No recalibration at all.
(2) A more flexible, monotone ("isotonic") recalibration instead of the linear
one — fitted with the same leave-one-study-out scheme. (3) Different cut-offs
for what counts as a miss.

```{r robustness, cache=TRUE}
iso_loo <- function(pred, est, study) {
  out <- rep(NA_real_, length(pred))
  for (s in unique(study)) {
    tr <- study != s
    fit <- isoreg(pred[tr], est[tr])
    f <- approxfun(sort(fit$x), fit$yf, method = "linear", rule = 2)
    out[!tr] <- f(pred[!tr])
  }
  out
}
zg_iso <- (iso_loo(wide$pg, wide$estimate, wide$study) - wide$estimate) / wide$std.error
zh_iso <- (iso_loo(wide$ph, wide$estimate, wide$study) - wide$estimate) / wide$std.error

summ <- function(a, b, thr = 1.96) {
  j <- abs(a) > thr & abs(b) > thr; s <- j & sign(a) == sign(b)
  nulls <- null_exact(mean(abs(a) > thr), mean(abs(b) > thr), t = thr)
  tibble(miss_g = mean(abs(a) > thr), miss_h = mean(abs(b) > thr),
         joint = mean(j), null_joint = nulls[["joint"]],
         same_side = mean(s), null_same = nulls[["same"]],
         direction = sum(s)/sum(j), null_direction = nulls[["direction"]])
}
bind_rows(
  `no recalibration`      = summ(wide$`z_raw_gpt-4`, wide$z_raw_human),
  `linear recalibration`  = summ(wide$zg, wide$zh),
  `isotonic recalibration`= summ(zg_iso, zh_iso),
  `linear, threshold 1.64`= summ(wide$zg, wide$zh, 1.64),
  `linear, threshold 2.58`= summ(wide$zg, wide$zh, 2.58),
  `linear, threshold 3.29`= summ(wide$zg, wide$zh, 3.29),
  .id = "analysis") %>% mutate(across(where(is.numeric), ~round(., 3))) %>% as.data.frame()

## the analysis set behind the 1,678, and a rank-based coupling measure
c(studies = n_distinct(wide$study),
  study_outcome_cells = wide %>% distinct(study, outcome.name) %>% nrow(),
  spearman_zg_zh = round(cor(wide$zg, wide$zh, method = "spearman"), 2))
```

```{r pp-corr}
er_g <- wide$zg * wide$std.error; er_h <- wide$zh * wide$std.error   # recalibrated errors, pp scale
c(recalibrated = round(cor(er_g, er_h), 2),
  uncorrected  = round(cor(wide$pg - wide$estimate, wide$ph - wide$estimate), 2))
```

# Does it hold under the authors' own sampling scheme?

**Why ask.** We analyse all 1,678 available comparisons. The authors instead
draw one outcome and one reference condition per study, compare every other
condition against that reference, and repeat the whole draw 32 times. A referee
will reasonably ask whether our result depends on using the full set.

**The check.** Reproduce their scheme exactly and recompute the two headline
rates, once with their 32 draws and once with many more draws so the spread is
stable.

```{r authors-sampling, cache=TRUE}
set.seed(22)
idx_by_study <- split(seq_len(nrow(wide)), wide$study)

one_draw <- function() {
  keep <- unlist(lapply(idx_by_study, function(ii) {
    ocs <- unique(wide$outcome.name[ii])
    oc  <- ocs[sample.int(length(ocs), 1)]              # one outcome per study
    jj  <- ii[wide$outcome.name[ii] == oc]
    cds <- unique(c(wide$reference_condition[jj], wide$condition.name[jj]))
    ref <- cds[sample.int(length(cds), 1)]              # one reference condition
    jj[wide$reference_condition[jj] == ref | wide$condition.name[jj] == ref]
  }), use.names = FALSE)
  g <- wide$zg[keep]; h <- wide$zh[keep]
  j <- abs(g) > 1.96 & abs(h) > 1.96
  c(n = length(keep), joint = mean(j), same = mean(j & sign(g) == sign(h)))
}

draws32  <- t(replicate(32,  one_draw()))
draws500 <- t(replicate(500, one_draw()))

rbind(
  `all 1,678 comparisons`       = c(NA, mean(joint), mean(same)),
  `their scheme, 32 draws`      = c(median(draws32[,  "n"]),
                                    median(draws32[,  "joint"]), median(draws32[,  "same"])),
  `their scheme, 500 draws`     = c(median(draws500[, "n"]),
                                    median(draws500[, "joint"]), median(draws500[, "same"]))
) %>% `colnames<-`(c("effects_per_draw", "joint", "same_side")) %>% round(3)

## spread across draws (a sensitivity check, not a confidence interval —
## the confidence intervals in the commentary come from the study bootstrap above)
apply(draws500[, c("joint", "same")], 2,
      function(x) c(sd = sd(x), lo = quantile(x, .025), hi = quantile(x, .975))) %>% round(3)
```

*Reading this:* under the authors' own sampling the joint-miss rate is about
23% and the same-side rate about 20% — both slightly **higher** than in the
full set (20.2% and 18.2%), and both well clear of their corrected nulls. The
finding is therefore not an artefact of analysing every available comparison.

Two notes. A draw yields about 235 unordered pairs, which is the authors'
reported 469 effects counted in both orientations: comparing every condition to
the reference produces one contrast per pair, and 305 conditions across 70
studies leaves 235 such pairs. The rates are unaffected by orientation anyway,
because flipping a pair flips the sign of the observed effect and of both
predictions together, leaving every |z| and every sign agreement unchanged.
Second, the spread across draws reported here measures sensitivity to their
sampling step; the uncertainty intervals quoted in the commentary continue to
come from the study-level bootstrap, which is the appropriate one.

# The figures as they appear in the commentary

Both main-text figures are drawn here, from the objects computed above, so the
numbers in the figures and the numbers in the text can never drift apart. The
chunks also save print-resolution PNGs next to this file.

```{r fig-style}
INK<-"#111111"; GREY<-"#5b5a57"; GRID<-"#e3e2dc"; RED<-"#c0392b"; BLUE<-"#1f6eb2"
PALE<-"#b9c9d9"; MUT<-"#c9c8c2"; GREEN<-"#1b8f63"; SAND<-"#d9a441"; FAINT<-"#dddcd6"
thm <- function(grid_y = TRUE, grid_x = FALSE) theme_minimal(base_size = 9) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.y = if (grid_y) element_line(colour = GRID, linewidth = .3) else element_blank(),
        panel.grid.major.x = if (grid_x) element_line(colour = GRID, linewidth = .3) else element_blank(),
        axis.text  = element_text(colour = GREY, size = 7.5),
        axis.title = element_text(colour = INK, size = 8),
        plot.title = element_text(colour = INK, size = 8.4),
        plot.tag   = element_text(face = "bold", size = 11),
        plot.margin = margin(4, 6, 4, 4))
```

```{r figure1, fig.width=7.2, fig.height=4.15, cache=TRUE}
## Panel a — the error plane
n_hi <- sum(wide$zg >  1.96 & wide$zh >  1.96)
n_lo <- sum(wide$zg < -1.96 & wide$zh < -1.96)
n_op <- sum(joint & !same)
pa <- wide %>%
  mutate(x = pmax(pmin(zg, 12), -12), y = pmax(pmin(zh, 12), -12),
         cls = case_when(same ~ "same", joint ~ "opp", TRUE ~ "rest")) %>%
  arrange(match(cls, c("rest", "opp", "same"))) %>%
  ggplot(aes(x, y)) +
  annotate("rect", xmin = 1.96, xmax = Inf, ymin = 1.96, ymax = Inf, fill = RED, alpha = .06) +
  annotate("rect", xmin = -Inf, xmax = -1.96, ymin = -Inf, ymax = -1.96, fill = RED, alpha = .06) +
  geom_hline(yintercept = 0, colour = "#9a9992", linewidth = .3) +
  geom_vline(xintercept = 0, colour = "#9a9992", linewidth = .3) +
  geom_hline(yintercept = c(-1.96, 1.96), colour = GREY, linetype = "22", linewidth = .3) +
  geom_vline(xintercept = c(-1.96, 1.96), colour = GREY, linetype = "22", linewidth = .3) +
  geom_point(aes(colour = cls, size = cls), alpha = .7, stroke = 0) +
  scale_colour_manual(values = c(same = RED, opp = "#4c7ba6", rest = MUT), guide = "none") +
  scale_size_manual(values = c(same = .9, opp = .9, rest = .55), guide = "none") +
  annotate("text", x = 12.6, y = 13.1, label = sprintf("both overshoot\n%d", n_hi),
           hjust = 1, vjust = 1, size = 2.6, colour = RED, lineheight = .95) +
  annotate("text", x = -12.6, y = -13.1, label = sprintf("both undershoot\n%d", n_lo),
           hjust = 0, vjust = 0, size = 2.6, colour = RED, lineheight = .95) +
  annotate("text", x = -12.6, y = 13.1, label = sprintf("opposite directions\n%d", n_op),
           hjust = 0, vjust = 1, size = 2.5, colour = "#4c7ba6", lineheight = .95) +
  scale_x_continuous(limits = c(-12.8, 12.8), breaks = seq(-10, 10, 5)) +
  scale_y_continuous(limits = c(-13.3, 13.3), breaks = seq(-10, 10, 5)) +
  labs(x = "GPT-4 error  (standard errors of the observed effect)",
       y = "Human crowd error  (standard errors)",
       title = "The two predictors' errors line up, comparison by comparison") +
  thm(grid_y = FALSE)

## Panel b — same-side rate against both nulls
b1 <- tibble(x = factor(c("Observed", "Corrected null", "Null ignoring\nshared estimate"),
                        levels = c("Observed", "Corrected null", "Null ignoring\nshared estimate")),
             v = 100 * c(mean(same), mean(same_s), mean(miss_g) * mean(miss_h) / 2),
             f = c(RED, MUT, FAINT))
pb <- ggplot(b1, aes(x, v)) +
  geom_col(fill = b1$f, width = .55) +
  geom_text(aes(label = sprintf("%.1f%%", v)), vjust = -.55, size = 2.9, fontface = "bold", colour = INK) +
  annotate("curve", x = 1.2, xend = 1.8, y = 100 * mean(same) + 1.5, yend = 100 * mean(same_s) + 2.5,
           curvature = -.35, colour = GREY, linewidth = .3) +
  annotate("text", x = 1.5, y = 25.4, label = sprintf("%.1f\u00d7", mean(same) / mean(same_s)),
           size = 2.9, fontface = "bold", colour = GREY) +
  scale_y_continuous(limits = c(0, 27), breaks = c(0, 10, 20), labels = c("0", "10%", "20%")) +
  labs(x = NULL, y = NULL, title = "Share of all 1,678 comparisons\nthat both predictors miss the same way") +
  thm() + theme(axis.text.x = element_text(colour = INK, size = 6.9, lineheight = .95))

## Panel c — direction given a joint miss
dirv <- 100 * sum(same) / sum(joint)
dir_null <- 100 * mean(sign(zgs[joint_s]) == sign(zhs[joint_s]))
c1 <- tibble(x = factor(c("Same\ndirection", "Opposite\ndirections"),
                        levels = c("Same\ndirection", "Opposite\ndirections")),
             v = c(dirv, 100 - dirv), f = c(RED, PALE))
pc <- ggplot(c1, aes(x, v)) +
  geom_col(fill = c1$f, width = .55) +
  geom_text(aes(label = sprintf("%.0f%%", v)), vjust = -.55, size = 2.9, fontface = "bold", colour = INK) +
  geom_hline(yintercept = dir_null, colour = GREY, linetype = "22", linewidth = .4) +
  annotate("text", x = 2.55, y = dir_null + 6.5, label = sprintf("corrected null, %.0f%%", dir_null),
           hjust = 1, size = 2.4, colour = GREY) +
  geom_hline(yintercept = 50, colour = "#b9b8b2", linetype = "12", linewidth = .35) +
  annotate("text", x = 2.55, y = 41, label = "chance, 50%", hjust = 1, size = 2.4, colour = "#9a9992") +
  scale_y_continuous(limits = c(0, 112), breaks = c(0, 50, 100), labels = c("0", "50%", "100%")) +
  labs(x = NULL, y = NULL, title = "Of the 339 comparisons both miss,\nhow the two misses line up") +
  thm() + theme(axis.text.x = element_text(colour = INK, size = 7.2, lineheight = .95))

fig1 <- pa + (pb / pc) + plot_layout(widths = c(1.52, 1)) +
  plot_annotation(tag_levels = "a") & theme(plot.tag.position = c(0, 1))
ggsave("F1_same_effects.png", fig1, width = 7.2, height = 4.15, dpi = 400, bg = "white")
fig1
```

```{r figure2, fig.width=7.2, fig.height=2.7, cache=TRUE}
## Panel a — every other predictor against GPT-4's misses
mods <- c(`deepseek/deepseek-chat-v3-0324` = "DeepSeek-V3", `openai/gpt-oss-120b` = "GPT-OSS-120B",
          `gpt-3.5-turbo` = "GPT-3.5", `google/gemma-3-27b-it` = "Gemma-3-27B",
          `babbage-002` = "babbage-002", `davinci-002` = "davinci-002", human = "Human crowd")
tab <- map_dfr(names(mods), function(mc) {
  zo <- zwide[[mc]]; ok <- !is.na(zo) & !is.na(gp); gm <- ok & abs(gp) > 1.96
  tibble(name = mods[[mc]],
         v = 100 * mean((abs(zo) > 1.96 & sign(zo) == sign(gp))[gm]),
         r = cor(gp[ok], zo[ok]))
}) %>% mutate(name = factor(name, levels = rev(name)),
              col  = ifelse(name == "Human crowd", GREEN, BLUE))
## the corrected-null benchmark for a predictor with the crowd's miss rate,
## reusing the simulated pair from the corrected-null chunk:
bench <- 100 * mean((abs(zhs) > 1.96 & sign(zhs) == sign(zgs))[abs(zgs) > 1.96])
p2a <- ggplot(tab, aes(v, name)) +
  geom_col(fill = tab$col, width = .62) +
  geom_text(aes(label = sprintf("%.0f%%", v)), hjust = -.25, size = 2.9, fontface = "bold", colour = INK) +
  geom_text(aes(x = 118, label = sprintf("r = %.2f", r)), hjust = 1, size = 2.5,
            fontface = "italic", colour = GREY) +
  geom_vline(xintercept = bench, colour = GREY, linetype = "22", linewidth = .4) +
  annotate("text", x = bench + 2, y = 8.9, label = sprintf("%.0f%% expected under the corrected null", bench),
           hjust = 0, vjust = 1, size = 2.35, colour = GREY) +
  scale_y_discrete(expand = expansion(add = c(.55, 1.7))) +
  scale_x_continuous(limits = c(0, 120), breaks = c(0, 25, 50, 75), labels = c("0", "25%", "50%", "75%")) +
  labs(x = "Also wrong, on the same side as GPT-4", y = NULL,
       title = "Every other predictor inherits GPT-4\u2019s misses") +
  coord_cartesian(clip = "off") +
  thm(grid_y = FALSE, grid_x = TRUE) +
  theme(axis.text.y = element_text(colour = INK, size = 7.6))

## Panel b — domain experts (second archive)
b2 <- tibble(x = factor(c("GPT-4\nmisses", "Experts\nmiss", "Experts also\nmiss, given\nGPT-4 misses"),
                        levels = c("GPT-4\nmisses", "Experts\nmiss", "Experts also\nmiss, given\nGPT-4 misses")),
             v = 100 * c(mean(gm2), mean(em2), sum(both2) / sum(gm2)),
             f = c(BLUE, SAND, RED))
p2b <- ggplot(b2, aes(x, v)) +
  geom_col(fill = b2$f, width = .55) +
  geom_text(aes(label = sprintf("%.0f%%", v)), vjust = -.55, size = 2.9, fontface = "bold", colour = INK) +
  annotate("text", x = 2.42, y = 112,
           label = sprintf("%.0f%% of those on\nthe same side", 100 * sum(same2) / sum(both2)),
           hjust = 1, size = 2.4, colour = RED, lineheight = .95) +
  scale_y_continuous(limits = c(0, 128), breaks = c(0, 50, 100), labels = c("0", "50%", "100%")) +
  labs(x = NULL, y = NULL, title = "Domain experts are not an\nindependent check") +
  thm() + theme(axis.text.x = element_text(colour = INK, size = 6.8, lineheight = .95))

fig2 <- p2a + p2b + plot_layout(widths = c(1.35, 1)) +
  plot_annotation(tag_levels = "a") & theme(plot.tag.position = c(0, 1))
ggsave("F2_shared_blind_spot.png", fig2, width = 7.2, height = 2.7, dpi = 400, bg = "white")
fig2
```

# How strongly are the two errors coupled, study by study?

```{r heterogeneity}
fit <- lm(zh ~ zg, data = wide)
## cluster-robust (by study) standard error, computed by hand:
X <- model.matrix(fit); u <- residuals(fit)
XtXi <- solve(crossprod(X))
meat <- Reduce(`+`, lapply(split(seq_len(nrow(X)), wide$study), function(i) {
  v <- crossprod(X[i, , drop = FALSE], u[i]); tcrossprod(v)
}))
se_cl <- sqrt(diag(XtXi %*% meat %*% XtXi))[2]
se_iid <- summary(fit)$coefficients[2, 2]
c(slope = round(coef(fit)[2], 2),
  cluster_ci_lo = round(coef(fit)[2] - 1.96 * se_cl, 2),
  cluster_ci_hi = round(coef(fit)[2] + 1.96 * se_cl, 2),
  iid_ci_lo = round(coef(fit)[2] - 1.96 * se_iid, 2),
  iid_ci_hi = round(coef(fit)[2] + 1.96 * se_iid, 2))

## share of the co-error's variance lying between studies (one-way ANOVA)
w <- wide$zg * wide$zh
gm <- mean(w)
ssb <- sum(tapply(w, wide$study, function(x) length(x) * (mean(x) - gm)^2))
round(100 * ssb / sum((w - gm)^2), 1)

slopes <- wide %>% group_by(study) %>% filter(n() >= 5, sd(zg) > 0) %>%
  summarise(slope = coef(lm(zh ~ zg))[2])
c(n_studies = nrow(slopes), median = round(median(slopes$slope), 2),
  q25 = round(quantile(slopes$slope, .25), 2), q75 = round(quantile(slopes$slope, .75), 2))
```

*Reading this:* on average, when GPT-4's error grows by one standard error, the
crowd's grows by about half of one (slope 0.46) — and the honest confidence
interval, which treats studies rather than comparisons as the unit, is
0.30–0.62. The per-study slopes show the coupling is everywhere but uneven.

# Session information

```{r session}
sessionInfo()
```
