---
title: "Supplementary Information"
subtitle: "Large language models and human forecasters make the same prediction errors"
author: "Live Leonhardsen Wilhelmsen, Roya Esfandiari & Anton Gollwitzer"
output:
  word_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressMessages({library(dplyr); library(knitr)})

## UTF-8 locale guard (see analysis_report.Rmd for why)
if (!l10n_info()[["UTF-8"]]) {
  for (loc in c("C.UTF-8", "C.utf8", "en_US.UTF-8"))
    if (nzchar(suppressWarnings(Sys.setlocale("LC_CTYPE", loc)))) break
}

## Locate the pipeline outputs relative to this document
find_file <- function(rel) {
  for (d in c("..", ".")) {
    p <- file.path(d, rel)
    if (file.exists(p)) return(p)
  }
  stop("Not found: ", rel, " — run `Rscript run_all.R --from-derived` first.")
}

rv   <- read.csv(find_file("output/tables/reported_values.csv"))
exc  <- read.csv(find_file("output/tables/exceedance_tests.csv"))
st1  <- read.csv(find_file("output/tables/supplementary_table_1.csv"), check.names = FALSE)
st2  <- read.csv(find_file("output/tables/supplementary_table_2.csv"), check.names = FALSE)
wide <- read.csv(find_file("data/derived/comparisons_primary.csv"),
                 check.names = FALSE, encoding = "UTF-8")
a2   <- read.csv(find_file("data/derived/comparisons_archive2.csv"),
                 check.names = FALSE, encoding = "UTF-8")

source(find_file("R/00_functions.R"))

## v(): fetch one canonical value; the knit FAILS if the value is absent,
## so this document cannot silently drift from the pipeline.
v <- function(loc, lab) {
  x <- rv$value[rv$location == loc & rv$label == lab]
  if (length(x) != 1) stop("reported_values.csv has no unique entry for [",
                           loc, " | ", lab, "]")
  x
}

pc   <- function(x, d = 1) sprintf(paste0("%.", d, "f%%"), 100 * x)  # 0.202 -> "20.2%"
pc0  <- function(x) sprintf("%.0f%%", 100 * x)                        # 0.90  -> "90%"
words <- function(n) if (n >= 1 && n <= 9)
  c("one","two","three","four","five","six","seven","eight","nine")[n] else n0(n)
n0   <- function(x) format(x, big.mark = ",")

## --- Quantities computed directly from the deposited tables -----------------

## archive sizes (Supplementary Methods 1.2)
n_comp    <- nrow(wide)
n_studies <- dplyr::n_distinct(wide$study)
n_conds   <- nrow(unique(rbind(
  data.frame(s = wide$study, c = wide$condition.name),
  data.frame(s = wide$study, c = wide$reference_condition))))
n_cells   <- nrow(unique(wide[, c("study", "outcome.name")]))

## pooled recalibration lines (Supplementary Methods 1.3)
fit_g <- lm(estimate ~ pg, data = wide)
fit_h <- lm(estimate ~ ph, data = wide)
int_g <- sprintf("%.3f", coef(fit_g)[1]); slo_g <- sprintf("%.2f", coef(fit_g)[2])
int_h <- sprintf("%.3f", coef(fit_h)[1]); slo_h <- sprintf("%.2f", coef(fit_h)[2])

## secondary archive (Supplementary Results 2.6)
a2e        <- a2[!is.na(a2$z_recal_expert) & !is.na(a2$z_recal_gpt), ]
n_a2_eff   <- nrow(a2)
n_a2_fam   <- dplyr::n_distinct(a2$family)
n_a2_exp   <- nrow(a2e)
n_a2_expfam <- dplyr::n_distinct(a2e$family)
gm2  <- abs(a2e$z_recal_gpt)    > 1.96
em2  <- abs(a2e$z_recal_expert) > 1.96
bt2  <- gm2 & em2
sm2  <- bt2 & (sign(a2e$z_recal_gpt) == sign(a2e$z_recal_expert))
null2       <- null_exact(mean(gm2), mean(em2))
naive_same2 <- mean(gm2) * mean(em2) / 2

## exceedance summaries (Supplementary Methods 1.7)
ex <- function(m, col) exc[[col]][exc$metric == m]
prange <- function(m, lo, hi, d = 1)
  paste0(pc(ex(m, lo), d), "–", pc(ex(m, hi), d))
pboot <- function(m, n_draws = 4000) {
  p <- ex(m, "p_bootstrap")
  if (is.na(p) || p == 0) paste0("*P* < ", format(1 / n_draws, scientific = FALSE))
  else if (p < 0.001) "*P* < 0.001"
  else paste0("*P* = ", sprintf("%.3f", p))
}
```

# 1 Supplementary Methods

## 1.1 Data

All data analysed were released with the original Article in Code Ocean capsule 9843791 [1]. Four files are used. rct_responses.RDS contains 233,601 individual responses from the experiments of the primary archive, of which 229,154 carry a non-missing response and enter the estimation. The raw archive comprises 71 studies and 134 study–outcome cells. llm_responses.RDS contains 504,840 language-model responses, 72,120 for each of seven models (GPT-4, GPT-3.5, davinci-002, babbage-002, DeepSeek-V3, Gemma-3-27B and GPT-OSS-120B), organized as 120 prompts per condition–outcome cell (three cells have 240), each prompt specifying a demographic profile and each recorded response the mean of five sampled completions. forecasting_responses.RDS contains 18,317 forecasts from 2,838 Prolific forecasters, a median of 29 per condition–outcome cell. Each forecaster predicted mean responses for a median of four conditions and, with three exceptions, within a single experiment. megastudies.RDS contains the secondary archive of `r n_a2_fam` megastudies: `r n0(n_a2_eff)` treatment effects, with expert forecasts available for `r n_a2_exp` conditions. The derivation of comparison-level data from this file is described in Supplementary Results 2.6. We recomputed all quantities from these files; none of the original data are redistributed with our code.

## 1.2 Comparisons and predictions

To construct the observed effects, we followed the authors' analytical pipeline. We first rescaled individual responses to range from 0 to 1 using the bounds of each outcome scale. Within each study and outcome, we then estimated every possible pairwise difference between conditions using ordinary least squares. For each comparison, we retained the estimated difference and its standard error. Reversing the order of two conditions changes only the sign of the estimated effect, and we therefore included each pair once. For reproducibility, we determined the direction of each comparison by sorting the condition names using a fixed English-language collation rule (ICU, en_US). This ordering matches the authors' deposited comparison-level dataset for every pair and does not materially affect the results because the recalibration intercepts are approximately zero (Supplementary Methods 1.3).

This procedure produced 1,705 possible comparisons. Our primary analysis includes the `r n0(n_comp)` comparisons where predictions from both GPT-4 and the human forecasters were available. These comparisons span `r n_studies` studies, `r n_conds` conditions and `r n_cells` study-outcome combinations. By contrast, the Article's headline analysis drew one reference condition from each study, yielding 469 treatment effects per draw. We apply the same sampling procedure to our measures as a robustness analysis in Supplementary Results 2.2.

We constructed predicted effects from the forecast condition means. For each language model, we calculated the predicted mean for a condition as a weighted average of its prompt-level responses, using the authors' demographic weights to account for their oversampling of population subgroups. For the human forecasters, we calculated an unweighted average of the individual forecasts. The predicted effect for each comparison is the difference between the predicted mean for its two conditions. Following the Article, the combined prediction is the arithmetic mean of the GPT-4 and human forecasters' predicted effects, formed before any recalibration.

## 1.3 Standardized errors and recalibration

Because the raw predictions systematically overstate effect magnitudes, we first recalibrate each predictor using the authors' leave-one-study-out linear procedure, which they used in their pilot. For each study, we regress observed effects on predicted effects using the `r n_studies - 1` studies and apply the resulting intercept and slope to the held-out study. We repeat this procedure separately for every study and predictor, including the combined predictor, which is therefore recalibrated in its own right rather than formed from two already-recalibrated predictions. This ensures that no study contributes to its own recalibration. The pooled recalibration lines have intercepts of `r int_g` for GPT-4 and `r int_h` for the human forecasters, with slopes of `r slo_g` and `r slo_h`, respectively. The leave-one-study-out estimates vary little around these values.

Using the recalibrated predictions, we express each prediction error in units of the observed effect's standard error. We classify a prediction as a miss when it differs from the observed effect by more than 1.96 standard errors, that is, when it falls outside the observed effect's 95% confidence interval. Recalibration reduces the overall miss rate from `r pc(v("Main text, para 7", "GPT-4 miss rate (unrecalibrated)"))` to `r pc(v("Main text, para 7", "GPT-4 miss rate (recalibrated)"))` for GPT-4 and from `r pc(v("Main text, para 7", "crowd miss rate (unrecalibrated)"))` to `r pc(v("Main text, para 7", "crowd miss rate (recalibrated)"))` for the human forecasters.

## 1.4 The corrected null

Some same-direction overlap is expected even when two predictors are independent. Both GPT-4 and the human forecasters are evaluated against the same estimated effect, which contains sampling error. If sampling error makes the observed effect unusually high, both predictions are more likely to appear too low. If it makes the effect unusually low, both are more likely to appear too high. A benchmark that ignores this shared source of error would therefore underestimate how often independent predictions should miss in the same direction.

We constructed a corrected null that treats the two predictors as independent while retaining the sampling error shared through the observed effect. We also adjusted the spread of each predictor's simulated errors to reproduce its observed miss rate. The benchmark therefore accounts for both the shared observed estimate and differences in the predictors' overall accuracy, while excluding any additional dependence between their errors.

At the primary threshold of 1.96, this null predicts a joint-miss rate of `r pc(v("Main text, para 9", "null joint rate"))`, a same-direction joint-miss rate of `r pc(v("Main text, para 9", "null same-side rate"))` and same-direction errors on `r pc(v("Main text, para 9", "null direction share"))` of joint misses. The corresponding observed values are `r pc(v("Main text, para 8", "joint miss rate"))`, `r pc(v("Main text, para 8", "same-side rate"))` and `r pc(v("Main text, para 8", "direction share given joint miss"))`. By comparison, a naive benchmark that ignores the shared observed estimate predicts a same-direction joint-miss rate of `r pc(v("Fig. 1b", "naive same-side null (ignores shared estimate)"))` and same-direction errors on 50% of joint misses (Fig. 1b).

We calculated the corrected benchmarks by numerical integration at each threshold used in our analyses and verified them using two million draws. Replacing the assumed normal error distribution with either a heavier-tailed distribution or the empirical error distribution produced slightly lower benchmarks (same-direction joint-miss rate `r sprintf("%.1f%%", v("SI 1.4", "same-side null, t(3)"))` under Student *t*(3) errors and `r sprintf("%.1f%%", v("SI 1.4", "same-side null, empirical"))` under the empirical distribution, against `r pc(v("Main text, para 9", "null same-side rate"))` under the normal). The normal specification reported in the main text is therefore the most conservative.

## 1.5 The study-specific null

The corrected null in Supplementary Methods 1.4 matches each predictor's miss rate across the full archive. However, studies differ in how difficult they are to predict. Even if GPT-4 and human errors were independent, both predictors would miss more often in difficult studies, producing greater overlap than the global benchmark expects.

We therefore constructed a stricter benchmark that matches the miss rates of GPT-4 and human forecasters separately within each study. As in the global null, the two predictors are otherwise treated as independent but remain exposed to the sampling error in the same observed effect. We generated 4,000 simulated replicates per comparison. When a predictor's miss rate within a study was at or below 5%, the rate attributable to sampling error alone, we assigned no additional predictor-specific error; this occurred in `r v("SI 1.5", "studies assigned zero spread")` studies and inflates the null in the conservative direction.

The study-specific benchmark takes differences in prediction difficulty across studies as given. Because the observed study-level miss rates may already contain some shared prediction error, the benchmark can absorb part of the dependence being tested. It is therefore more conservative than the global null.

Under the study-specific null, the expected joint-miss rate is `r pc(v("SI 1.5", "study-specific null: joint rate"))`, the expected same-direction joint-miss rate is `r pc(v("SI 1.5", "study-specific null: same-side rate"))` and `r pc(v("SI 1.5", "study-specific null: direction share"))` of joint misses are expected to occur in the same direction. The corresponding observed values are `r pc(v("Main text, para 8", "joint miss rate"))`, `r pc(v("Main text, para 8", "same-side rate"))` and `r pc(v("Main text, para 8", "direction share given joint miss"))`.

## 1.6 Uncertainty intervals

Comparisons from the same study are not independent because they share participants and often a reference condition. We therefore estimate uncertainty by resampling studies rather than individual comparisons. The 95% confidence intervals are percentile intervals from a nonparametric cluster bootstrap [2] that resamples the `r n_studies` studies with replacement 4,000 times. The resulting intervals are `r pc(v("SI 1.6", "joint rate 95% CI lower"))`–`r pc(v("SI 1.6", "joint rate 95% CI upper"))` for the joint-miss rate, `r pc(v("Main text, para 8", "same-side rate 95% CI lower"))`–`r pc(v("Main text, para 8", "same-side rate 95% CI upper"))` for the same-direction joint-miss rate and `r pc(v("Main text, para 8", "direction share 95% CI lower"))`–`r pc(v("Main text, para 8", "direction share 95% CI upper"))` for the proportion of joint misses occurring in the same direction.

The error bars in the figures follow the same logic. Observed archive-wide rates carry the study-level bootstrap interval described above. Corrected-null bars carry the 95% range of the finite-sample Monte Carlo distribution described in Supplementary Methods 1.7, and the bar for the naive benchmark that ignores the shared observed estimate carries a binomial interval at the same archive size. Conditional shares — the proportion of GPT-4's misses that another predictor repeats, and the proportions in the secondary archive, where only `r n_a2_fam` megastudy families (`r words(n_a2_expfam)` with expert forecasts) are available for resampling — carry Wilson 95% intervals.

The shaded band around the fitted line in Fig. 1a is the 95% confidence band of the regression of the human forecasters' standardized errors on GPT-4's, with standard errors clustered by study (Supplementary Results 2.3).

## 1.7 Exceedance tests

The corrected null gives the rates that independent predictors would be expected to produce; it does not by itself establish whether the observed rates fall outside what sampling variation under that null could generate. We therefore tested the observed joint-miss rate, same-direction joint-miss rate and direction share against the null benchmarks in three complementary ways. All tests are one-sided.

First, we simulated 10,000 archives of `r n0(n_comp)` comparisons from the global corrected null, using the same calibration as in Supplementary Methods 1.4. The *P* value is the share of simulated archives whose rate reaches the observed one. Under this null, the 95% ranges are `r prange("joint", "null_global_lo", "null_global_hi")` for the joint-miss rate, `r prange("same_side", "null_global_lo", "null_global_hi")` for the same-direction joint-miss rate and `r prange("direction", "null_global_lo", "null_global_hi")` for the proportion of joint misses occurring in the same direction. None of the 10,000 archives reached the observed values of `r pc(v("Main text, para 8", "joint miss rate"))`, `r pc(v("Main text, para 8", "same-side rate"))` and `r pc(v("Main text, para 8", "direction share given joint miss"))` (*P* < 0.0001 for each).

Second, we repeated the test against the stricter study-specific null of Supplementary Methods 1.5, simulating 2,000 archives with the spreads calibrated within each study. The corresponding 95% ranges are `r prange("joint", "null_ss_lo", "null_ss_hi")`, `r prange("same_side", "null_ss_lo", "null_ss_hi")` and `r prange("direction", "null_ss_lo", "null_ss_hi")`; again no archive reached the observed values (*P* < 0.001 for each).

Third, because both Monte Carlo tests draw comparisons independently whereas real comparisons cluster within studies, we inverted the study-level bootstrap of Supplementary Methods 1.6 into a test: the *P* value is the share of the 4,000 bootstrap draws whose rate falls at or below the corrected-null benchmark. This gives `r pboot("joint")` for the joint-miss rate and `r pboot("same_side")` for the same-direction joint-miss rate; no draw fell at or below the null for the direction share (`r pboot("direction")`). All three approaches lead to the same conclusion: the observed overlap exceeds what independent predictors sharing only the observed estimate would produce.

# 2 Supplementary Results

## 2.1 Robustness to recalibration and threshold

Supplementary Table 1 shows that the shared-error pattern is robust to the recalibration method and threshold used. The observed joint-miss and same-direction miss rates exceed their corrected-null benchmarks without recalibration, under linear and isotonic recalibration, and at thresholds ranging from 1.64 to 3.29. At stricter thresholds, misses become less frequent but exceed the null by larger relative margins. On the original percentage-point scale, GPT-4 and human forecaster errors correlate at *r* = `r sprintf("%.2f", v("SI 2.1", "error correlation, recalibrated (pp metric)"))` after linear recalibration and *r* = `r sprintf("%.2f", v("SI 2.1", "error correlation, unrecalibrated (pp metric)"))` without recalibration.

```{r supp-table-1}
st1 %>%
  transmute(
    Analysis      = analysis,
    `GPT-4 miss`  = pc(miss_g),
    `Forecaster miss` = pc(miss_h),
    `Joint (null)`     = paste0(pc(joint), " (", pc(null_joint), ")"),
    `Same side (null)` = paste0(pc(same_side), " (", pc(null_same), ")"),
    `Direction (null)` = paste0(pc(direction), " (", pc(null_direction), ")")) %>%
  kable(align = c("l", rep("r", 5)),
        caption = paste("Supplementary Table 1 | Shared error under alternative",
                        "recalibrations and thresholds. Values in parentheses are",
                        "corrected-null benchmarks calculated separately for each row.",
                        '"Same side" is the proportion of all comparisons missed by both',
                        'predictors in the same direction; "Direction" is the proportion',
                        "of joint misses occurring in the same direction. Isotonic",
                        "recalibration uses a monotonic fit under the same",
                        "leave-one-study-out procedure, with predictions outside the",
                        "training range held at the nearest boundary."))
```

## 2.2 Study weighting and the authors' sampling scheme

The `r n_studies` studies contribute between `r v("SI 2.2", "smallest study, comparisons")` and `r v("SI 2.2", "largest study, comparisons")` comparisons each, so the main-text estimates give greater weight to studies with more comparisons. When each study is weighted equally, the joint-miss rate increases from `r pc(v("Main text, para 8", "joint miss rate"))` to `r pc(v("SI 2.2", "study-weighted joint rate"))` and the same-direction joint-miss rate increases from `r pc(v("Main text, para 8", "same-side rate"))` to `r pc(v("SI 2.2", "study-weighted same-side rate"))`.

We also applied the Article's own sampling procedure. The Article's Supplementary Information describes the draw as selecting one reference condition and one dependent variable per study, but that selection would retain 235 contrasts (`r n_conds` conditions minus `r n_studies` studies), not the 469 effects the Article reports. Drawing one reference condition per study and keeping all outcomes retains exactly `r v("SI 2.2", "comparisons retained per draw (median)")` comparisons on every draw, matching the Article's count, and is the scheme we reproduce. Across 500 draws, the joint-miss and same-direction rates had medians of `r pc(v("SI 2.2", "authors' scheme, 500 draws: joint"))` and `r pc(v("SI 2.2", "authors' scheme, 500 draws: same-side"))`, with a standard deviation of approximately `r sprintf("%.1f", 100 * v("SI 2.2", "s.d. across draws, same-side"))` percentage points across draws. This variation is a sensitivity check, not a confidence interval. The confidence intervals reported in the main text come from the study-level bootstrap described in Supplementary Methods 1.6.

## 2.3 Study-level heterogeneity

The shared errors were not confined to a small number of studies. The `r v("Main text, para 8", "joint misses (n)")` joint misses occurred across `r v("SI 2.3", "studies with a joint miss")` of the `r n_studies` studies, and the `r v("Main text, para 8", "same-side misses (n)")` same-direction joint misses across `r v("SI 2.3", "studies with a same-side miss")`. Across the full archive, regressing the human forecasters' standardized errors on GPT-4's produced a slope of `r sprintf("%.2f", v("SI 2.3", "pooled slope of crowd error on GPT-4 error"))` (95% CI `r sprintf("%.2f", v("SI 2.3", "slope CI lower (clustered by study)"))`–`r sprintf("%.2f", v("SI 2.3", "slope CI upper (clustered by study)"))`, with standard errors clustered by study). The strength of this association varied across studies. Among the `r v("SI 2.3", "studies with at least five comparisons")` studies containing at least five comparisons, the median study-specific slope was `r sprintf("%.2f", v("SI 2.3", "median per-study slope"))` (interquartile range, `r sprintf("%.2f", v("SI 2.3", "per-study slope, 25th percentile"))`–`r sprintf("%.2f", v("SI 2.3", "per-study slope, 75th percentile"))`). Study membership accounted for `r sprintf("%.1f%%", v("SI 2.3", "percent of co-error variance between studies"))` of the variance in the co-error — the product of the two predictors' standardized errors — further indicating meaningful between-study variation.

## 2.4 The continuous co-error

To test whether the findings depend on classifying predictions as misses, we also used a continuous co-error measure. This measure multiplies the two standardized errors and is positive when they point in the same direction. Its observed mean was `r sprintf("%.2f", v("SI 2.4", "observed mean co-error"))`, compared with 1 under the corrected null. Randomly breaking the pairing between GPT-4 and human forecaster errors produced a mean of `r sprintf("%.2f", abs(v("SI 2.4", "permutation mean co-error")))` across 2,000 permutations (s.d. = `r sprintf("%.2f", v("SI 2.4", "permutation s.d."))`). The co-error was positive on average in `r v("SI 2.4", "studies with positive mean co-error")` of the `r v("SI 2.4", "studies total")` studies. A Wilcoxon signed-rank test of the study-level mean co-errors against zero gives *P* = `r sub("e-0?", " × 10^-", sprintf("%.1e", v("SI 2.4", "Wilcoxon P")))`^, treating the study rather than the comparison as the unit of analysis. The predictors' standardized errors also had a Spearman correlation of `r sprintf("%.2f", v("SI 2.1", "Spearman correlation of standardized errors"))`, indicating that the association was not driven by extreme errors.

## 2.5 Dependence across language models

Shared errors with GPT-4 extend to every other predictor in the archive. Among comparisons missed by GPT-4, the other language models miss the same comparison in the same direction in `r pc0(min(st2$also_same_side[st2$predictor != "human"]))`–`r pc0(max(st2$also_same_side[st2$predictor != "human"]))` of cases, compared with `r pc0(min(st2$corrected_null[st2$predictor != "human"]))`–`r pc0(max(st2$corrected_null[st2$predictor != "human"]))` under the corrected null. The human forecasters have the lowest overlap with GPT-4, at `r pc0(st2$also_same_side[st2$predictor == "human"])` compared with a corrected-null benchmark of `r pc0(st2$corrected_null[st2$predictor == "human"])`.

```{r supp-table-2}
pretty <- c("deepseek/deepseek-chat-v3-0324" = "DeepSeek-V3",
            "openai/gpt-oss-120b"            = "GPT-OSS-120B",
            "gpt-3.5-turbo"                   = "GPT-3.5",
            "google/gemma-3-27b-it"           = "Gemma-3-27B",
            "babbage-002"                     = "babbage-002",
            "davinci-002"                     = "davinci-002",
            "human"                           = "Human forecasters")
st2 %>%
  transmute(
    Predictor = pretty[predictor],
    `Miss rate` = pc0(miss_rate),
    `Also misses, same side` = pc0(also_same_side),
    `Corrected null` = pc0(corrected_null),
    `Error correlation` = sprintf("%.2f", error_corr)) %>%
  kable(align = c("l", rep("r", 4)),
        caption = paste("Supplementary Table 2 | Shared misses between GPT-4 and each",
                        "other predictor in the archive. All predictors were recalibrated",
                        "using the same leave-one-study-out linear procedure. Miss rates",
                        "are calculated on comparisons for which predictions from both",
                        "GPT-4 and the relevant predictor are available. Corrected-null",
                        "values represent the expected same-direction overlap with GPT-4",
                        "for an independent predictor with the same miss rate",
                        "(Supplementary Methods 1.4). Error correlations are Pearson",
                        "correlations between the predictors' standardized errors."))
```

## 2.6 Secondary archive of megastudies

The secondary archive contains `r n0(n_a2_eff)` treatment effects from `r n_a2_fam` megastudies, with expert forecasts available for `r n_a2_exp` conditions in `r words(n_a2_expfam)` of them. We derived the comparison-level data from the authors' megastudies.RDS file, following their scale conversions, family-specific scale ratios and data corrections. Because outcome scales differ across megastudy families, we recalibrated predictions separately within each family using a leave-one-condition-out procedure.

Across the `r n_a2_exp` conditions, experts missed `r pc(mean(em2))` of the observed effects and GPT-4 missed `r pc(mean(gm2))`. Of the `r sum(gm2)` conditions missed by GPT-4, experts also missed `r sum(bt2)` (`r pc(sum(bt2) / sum(gm2))`); `r sum(sm2)` of these `r sum(bt2)` joint misses (`r pc(sum(sm2) / sum(bt2))`) were in the same direction. Both predictors therefore missed `r sum(bt2)` of the `r n_a2_exp` conditions (`r pc(mean(bt2))`), and missed in the same direction on `r sum(sm2)` (`r pc(mean(sm2))` of all conditions). Under the corrected null, the expected same-direction joint-miss rate is `r pc(null2[["same"]])`; under the naive benchmark that ignores the shared observed estimate, it is `r pc(naive_same2)`. Under the corrected null, experts would be expected to miss `r pc(v("SI 2.6", "null: experts also miss | GPT-4 misses"))` of GPT-4's misses, with `r pc(v("SI 2.6", "null: same side | both miss"))` of joint misses occurring in the same direction. Expert errors also closely tracked GPT-4's errors, with a regression slope of `r sprintf("%.2f", v("Main text, para 12", "slope of expert error on model error"))` (95% CI `r sprintf("%.2f", v("Main text, para 12 (C17)", "expert slope CI lower (clustered)"))`–`r sprintf("%.2f", v("Main text, para 12 (C17)", "expert slope CI upper (clustered)"))`, with standard errors clustered by megastudy family). Because the analysis contains only `r words(n_a2_expfam)` megastudy-family clusters, this interval should be interpreted with caution.

## 2.7 Concentration of same-side joint misses

Same-direction joint misses were more common for larger observed effects. Using the Article's five-percentage-point threshold, rates were `r pc(v("Main text, para 14", "same-side rate | larger effects"))` across `r v("SI 2.7", "comparisons with larger observed effects")` larger-effect comparisons and `r pc(v("Main text, para 14", "same-side rate | smaller effects"))` across `r v("SI 2.7", "comparisons with smaller observed effects")` smaller-effect comparisons. Some difference is expected because grouping by a noisy observed estimate selects large effects inflated by sampling error, while both forecasts remain more moderate. Accordingly, the corrected-null benchmarks were `r pc(v("SI 2.7", "corrected null, larger effects"))` and `r pc(v("SI 2.7", "corrected null, smaller effects"))`, respectively. The observed rates exceeded both benchmarks, indicating that sampling error partly — but not fully — explains the concentration. We therefore do not rely on this split in the main text.

We next defined the profile of predictions a researcher would be most likely to treat as compelling evidence: both predictors expecting a large effect (top fifth by the magnitude of the average raw prediction), agreeing in sign, and closely aligned in magnitude (disagreement below the median). `r v("Main text, para 14", "screening cell size")` comparisons met these criteria, of which `r pc(v("Main text, para 14", "same-side rate inside screening cell"))` were same-direction joint misses, compared with `r pc(v("SI 2.7", "same-side rate outside screening cell"))` of the remaining comparisons and `r pc(v("SI 2.7", "corrected null, screening cell"))` expected under the corrected null. Because the profile is based on predictions rather than observed effects, it avoids the selection issue above and reflects how researchers might screen on LLM–human forecaster agreement. We therefore report this analysis in the main text.

This analysis describes where shared misses are concentrated; it does not independently establish that the two predictors' errors are dependent. Because both predictions are evaluated against the same noisy observed estimate, sampling error alone can place them on the same side. Dependence is instead evaluated using the corrected nulls in Supplementary Methods 1.4 and 1.5.

## 2.8 Bounding the share of misses attributable to inaccurate estimates

An observed effect may itself be inaccurate, in which case a joint miss could reflect sound forecasts and a faulty estimate rather than shared forecasting error. This section bounds how much of the same-direction overlap that explanation can carry. If an estimate's *t*-interval contains the true effect, a forecast outside that interval is necessarily different from the truth, so the miss cannot be attributed to the estimate alone. Under calibrated standard errors the true effect lies outside the *t*-interval with probability 2(1 − Φ(*t*)), which is `r pc(v("SI 2.8", "threshold 1.96: calibration cap"))` at *t* = 1.96 and `r pc(v("SI 2.8", "threshold 3.29: calibration cap"))` at *t* = 3.29. The share of comparisons whose same-direction joint miss the estimate alone could explain is therefore capped at those rates.

The observed same-direction joint-miss rates are `r pc(v("SI 2.8", "threshold 1.96: same-side joint-miss rate"))` and `r pc(v("SI 2.8", "threshold 3.29: same-side joint-miss rate"))`, exceeding the caps by factors of `r sprintf("%.1f", v("SI 2.8", "threshold 1.96: observed / cap"))` and `r sprintf("%.0f", v("SI 2.8", "threshold 3.29: observed / cap"))`. In expectation, at least `r pc(v("SI 2.8", "threshold 1.96: floor share, expectation"))` of same-direction joint misses at the 1.96 threshold, and at least `r pc(v("SI 2.8", "threshold 3.29: floor share, expectation"))` at the 3.29 threshold, must therefore occur on comparisons whose interval contains the true effect, where both forecasts are demonstrably in error.

The cap is an expected rate, and the realized number of intervals excluding the truth varies from archive to archive. As sensitivity checks we recompute each floor with a binomial 97.5% upper limit on that number, and with intervals assumed to fail at ten times their nominal rate. Because comparisons cluster within studies, the binomial limit is indicative rather than exact. At the 1.96 threshold the floor is `r pc(v("SI 2.8", "threshold 1.96: floor share, binomial 97.5% cap"))` under the binomial limit but does not survive substantial miscalibration. At the 3.29 threshold it is `r pc(v("SI 2.8", "threshold 3.29: floor share, binomial 97.5% cap"))` under the binomial limit and `r pc(v("SI 2.8", "threshold 3.29: floor share, tenfold miscalibration"))` under tenfold miscalibration. The claim carried into the main text therefore rests on the strictest threshold, where it is robust. Calibration itself is defended by the archive's design. The experiments are pre-registered, every pairwise comparison is analysed rather than only the statistically significant ones, and the estimates are simple mean differences from large samples, so the selection mechanisms that undermine interval calibration in the published literature are absent by construction.

# 3 Data and code provenance

All analyses were produced in R 4.3, starting from the raw files in Code Ocean capsule 9843791 [1]. Reanalysis code, the deposited comparison-level tables and a self-verifying analysis report are available at https://github.com/Livedenkule/Large-language-models-and-human-crowds-make-the-same-prediction-errors- (archived at [ZENODO DOI ON PUBLICATION]). The report re-computes every quantity in this document and refuses to render if any reported value fails to reproduce.

# Supplementary references

[1] Ashokkumar, A., Hewitt, L., Ghezae, I. & Willer, R. Data and code for: Large language models can predict the results of social science experiments. Code Ocean https://codeocean.com/capsule/9843791/tree/v1 (2026).

[2] Efron, B. & Tibshirani, R. J. An Introduction to the Bootstrap (Chapman & Hall, 1993).
